---
title: "Sickle_Cell_Project"
author: "Sedat Kacar"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(Seurat)
library(dplyr)
library(ggplot2)
```

VPC1 = WT Male 
VPC2  = Sickle Male
VPC3 = WT Female 
VPC4 = Sickle Female 

```{r}

```


```{r}
Vpc1 <- Read10X_h5(file = "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/1.Data_process/VPC1/WT_Male_filtered_feature_bc_matrix.h5")
Vpc2 <- Read10X_h5(file = "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/1.Data_process/VPC2/Sickle_Male_filtered_feature_bc_matrix.h5")
Vpc3 <- Read10X_h5(file = "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/1.Data_process/VPC3/WT_Female_filtered_feature_bc_matrix.h5")
Vpc4 <- Read10X_h5(file = "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/1.Data_process/VPC4/Sickle_Female_filtered_feature_bc_matrix.h5")

```


```{r}
Vpc1 <- CreateSeuratObject(Vpc1,
                                    min.cells=3,
                                    min.features = 100)

Vpc2 <- CreateSeuratObject(Vpc2,
                                    min.cells=3,
                                    min.features = 100)

Vpc3 <- CreateSeuratObject(Vpc3,
                                    min.cells=3,
                                    min.features = 100)
Vpc4 <- CreateSeuratObject(Vpc4,
                                    min.cells=3,
                                    min.features = 100)


rownames(Vpc1)[grep("^mt-", rownames(Vpc1))]
```


```{r}

for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
  obj <- get(i)
  obj[["percent.mt"]] <- PercentageFeatureSet(obj, pattern = "^mt-")
  assign(i, obj)
}

```



```{r, fig.height=15, fig.width=10}
library(ggplot2)

for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
obj <- get(i)
# Plot features vs counts
p1 <- FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_hline(yintercept = 5000, linetype = "dashed", color = "red", size = 1.5) + geom_vline(xintercept = c(800, 17000), linetype = "dashed", color = "red", size = 1.5)  + theme(legend.position = "none") + ggtitle(i)
# Display the combined plot
p2 <- VlnPlot(obj, features = "percent.mt") + geom_hline(yintercept = 10, linetype = "dashed", color = "red", size = 1.5) + theme(legend.position = "none") + ggtitle(i)
print(p1/p2)
}

```



```{r}
for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
  obj <- get(i)
obj <- subset(obj, subset = nFeature_RNA > 100 & nFeature_RNA < 5000 & nCount_RNA < 17000 & nCount_RNA > 800 & percent.mt < 10)
assign(i, obj)
}
```


```{r}
for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj, selection.method = "vst", nfeatures = 2000)
assign(i, obj)
}
```


```{r}
library(Seurat)

```


```{r}
for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
obj$orig.ident <- i
assign(i, obj)
}
```

```{r}
sickle_list <- list(WT_Male = Vpc1, Sickle_Male = Vpc2, WT_Female = Vpc3, Sickle_Female = Vpc4)
```

```{r}

anchors <- FindIntegrationAnchors(object.list = sickle_list)

```

```{r}
sickle_cell <- IntegrateData(anchorset = anchors)

```
#Vpc1
```{r, fig.height=12, fig.width=10}

for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
obj <- ScaleData(obj)
obj <- RunPCA(obj, npcs = 50)
obj <- RunUMAP(obj, dims = 1:50)
obj <- FindNeighbors(obj, dims = 1:30)
obj <- FindClusters(obj, resolution = 0.30)
assign(i, obj)
}


sickle_list <- list(WT_Male = Vpc1, Sickle_Male = Vpc2, WT_Female = Vpc3, Sickle_Female = Vpc4)

```

#RNA

```{r, fig.height=12, fig.width=10}

DefaultAssay(sickle_cell) <- "RNA"
sickle_cell <- ScaleData(sickle_cell, rownames(sickle_cell))

sickle_cell <- RunPCA(sickle_cell, npcs = 50)
sickle_cell <- RunUMAP(sickle_cell, dims = 1:50)
sickle_cell <- FindNeighbors(sickle_cell, dims = 1:30)
sickle_cell <- FindClusters(sickle_cell, resolution = 0.30)
DimPlot(sickle_cell, label = T) + theme(legend.position = "bottom")
DimPlot(sickle_cell, label = T, split.by = "orig.ident") + theme(legend.position = "bottom")

library(gridExtra)
library(ggplot2) 
# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("Seurat_sickle_0.30_new.png", width = 9, height = 9, units = "in", res = 300)

# Arrange and save all plots in a grid with 3 columns
DimPlot(sickle_cell, label = T) + theme(legend.position = "bottom") + ggtitle("Sickle Cell Project")

# Close the graphics device
dev.off()

```



```{r, fig.height=12, fig.width=10}
#Constructing the group name with ifelse
sickle_cell$group <- ifelse(grepl("_1$", colnames(sickle_cell)), "WT Male",
                               ifelse(grepl("_2$", colnames(sickle_cell)), "ScAnemia Male",
                                      ifelse(grepl("_3$", colnames(sickle_cell)), "WT Female", "ScAnemia Female")))
```

#TRANSFER DATA WITH PARSE

```{r, fig.height=10, fig.width=10}
#BISMILLAHIRRAHMANIRRAHIM

Parsall_comb$general_group_nounk
Parsall_comb$general_group_w_na

Vpc1 <- sickle_list[1]$Vpc1
Vpc2 <- sickle_list[2]$Vpc2
Vpc3 <- sickle_list[3]$Vpc3
Vpc4 <- sickle_list[4]$Vpc4

for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
    anchors <- FindTransferAnchors(
      reference = Parsall_comb,         # Reference Seurat object (with known T cell labels)
  query = obj,                 # Query Seurat object (to which we want to transfer labels)
  normalization.method = "LogNormalize",       # Normalization method (use "SCT" if you did SCTransform or "SCT")
  reference.assay = "RNA",            # Assay in the reference Seurat object (usually "RNA" for raw or normalized counts)
  query.assay = "RNA",                # Assay in the query Seurat object
  dims = 1:30)
    #2- Second Tranfer labels
    Bis_TRANSITIVE_Elh <- TransferData(anchorset = anchors,
                             refdata = Parsall_comb$general_group_nounk,
                             dims = 1:30
                             )
    obj$general_group_nounk <- Bis_TRANSITIVE_Elh$predicted.id
    assign(i, obj)
}

sickle_cell$general_group_nounk <- NA


for (i in 1:4){
    colnames(sickle_list[[i]]) <- paste0(colnames(sickle_list[[i]]),"_", i)
}



for (i in 1:4){
      obj <- sickle_list[[i]]
      yes <- colnames(obj)
      sickle_cell$general_group_nounk[yes] <- obj$general_group_nounk[yes]
}

   p <- DimPlot(sickle_cell, group.by = "general_group_nounk", label = T) + 
      theme(legend.position = "bottom") + ggtitle(i)
      LabelClusters(p, id = "general_group_nounk", fontface = "bold", size = 5)
      print(p)




#SEE GENERAL GROUP IN DATAFRAME-START
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
df1 <- as.data.frame(table(Vpc1$general_group_nounk))
df2 <-  as.data.frame(table(Vpc2$general_group_nounk))
df3 <- as.data.frame(table(Vpc3$general_group_nounk))
df4 <- as.data.frame(table(Vpc4$general_group_nounk))

rownames(df1) <- df1$Var1
rownames(df2) <- df2$Var1
rownames(df3) <- df3$Var1
rownames(df4) <- df4$Var1

df$Var1 <- NULL
df1$Var1 <- NULL
df2$Var1 <- NULL
df3$Var1 <- NULL
df4$Var1 <- NULL

colnames(df1)[1] <- "WT Male"
colnames(df2)[1] <- "ScAnemia Male"
colnames(df3)[1] <- "WT Female"
colnames(df4)[1] <- "ScAnemia Female"

df <- cbind(df1, df2, df3, df4)


library(ggplot2)
library(tidyr)

df$Cell_Type <- rownames(df)

# Convert data frame to long format
df_long <- pivot_longer(df, cols = c("WT Male", "ScAnemia Male", "WT Female", "ScAnemia Female"),
                        names_to = "Condition", values_to = "Count")


library(gridExtra)
library(ggplot2) 
# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("General_group_sickle_number.png", width = 5, height = 6, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()



# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = Count, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")



df_long$total <- NA
df_long$total[df_long$Condition == "WT Male"] <- 7973
df_long$total[df_long$Condition == "ScAnemia Male"] <- 7365
df_long$total[df_long$Condition == "WT Female"] <- 8886
df_long$total[df_long$Condition == "ScAnemia Female"] <- 7378


df_long$count_percent <- NA
df_long$count_percent <- (df_long$Count * 100) / df_long$total





p <- ggplot(df_long, aes(x = Cell_Type, y = count_percent, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f", count_percent)), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Percentage Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

png("General_group_sickle_percentage.png", width = 5, height = 6, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()




#SEE GENERAL GROUP IN DATAFRAME-END
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------

Parsall_comb$Annotation_transfer_EC
Parsall_comb$Marker_Ann

columns <- c("Annotation_transfer_EC", "Marker_Ann")

for (i in columns){

sickle_cell[[i]] <- NA
}



for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
    anchors <- FindTransferAnchors(
      reference = Parsall_comb,         # Reference Seurat object (with known T cell labels)
  query = obj,                 # Query Seurat object (to which we want to transfer labels)
  normalization.method = "LogNormalize",       # Normalization method (use "SCT" if you did SCTransform or "SCT")
  reference.assay = "RNA",            # Assay in the reference Seurat object (usually "RNA" for raw or normalized counts)
  query.assay = "RNA",                # Assay in the query Seurat object
  dims = 1:30)
                #2- Second Tranfer labels
    #first label
    Bis_TRANSITIVE_Elh <- TransferData(anchorset = anchors,
                             refdata = Parsall_comb$Annotation_transfer_EC,
                             dims = 1:30
                             )
    obj$Annotation_transfer_EC<- Bis_TRANSITIVE_Elh$predicted.id
    yes <- colnames(obj)
    sickle_cell$Annotation_transfer_EC[yes] <- obj$Annotation_transfer_EC[yes]
  
    #Second label
      Bis_TRANSITIVE_Elh <- TransferData(anchorset = anchors,
                             refdata = Parsall_comb$Marker_Ann,
                             dims = 1:30
                             )
    obj$Marker_Ann<- Bis_TRANSITIVE_Elh$predicted.id
    yes <- colnames(obj)
    sickle_cell$Marker_Ann[yes] <- obj$Marker_Ann[yes]
    assign(i, obj) 
}


for (i in 1:4){

sickle_list[[i]]$Annotation_transfer_EC <- NA
sickle_list[[i]]$Marker_Ann <- NA
}

    colnames(Vpc1) <- paste0(colnames(Vpc1),"_", 1)
    colnames(Vpc2) <- paste0(colnames(Vpc2),"_", 2)
    colnames(Vpc3) <- paste0(colnames(Vpc3),"_", 3)
    colnames(Vpc4) <- paste0(colnames(Vpc4),"_", 4)


sickle_list[[1]]$Annotation_transfer_EC <- Vpc1$Annotation_transfer_EC
sickle_list[[2]]$Annotation_transfer_EC<- Vpc2$Annotation_transfer_EC
sickle_list[[3]]$Annotation_transfer_EC<- Vpc3$Annotation_transfer_EC
sickle_list[[4]]$Annotation_transfer_EC<- Vpc4$Annotation_transfer_EC



sickle_list[[1]]$Marker_Ann <- Vpc1$Marker_Ann
sickle_list[[2]]$Marker_Ann<- Vpc2$Marker_Ann
sickle_list[[3]]$Marker_Ann<- Vpc3$Marker_Ann
sickle_list[[4]]$Marker_Ann<- Vpc4$Marker_Ann

sickle_cell$general_group_nounk <- NA

  yes <- colnames(Vpc1)


for(i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
  obj <- get(i)
  yes <- colnames(obj)
  sickle_cell$Annotation_transfer_EC[yes] <- obj$Annotation_transfer_EC[yes]
         sickle_cell$Marker_Ann[yes]<- obj$Marker_Ann[yes] 

}

  
```
  
  
 # UMAP GENERAL 
  
```{r, fig.height=19, fig.width=14}

pt_size <- 0.7

p1 <- DimPlot(sickle_cell, group.by = "Annotation_transfer_EC", label = F, pt.size = pt_size) + 
  theme(legend.position = "bottom") + ggtitle("Annotation_transfer_EC")
p1 <- LabelClusters(p1, id = "Annotation_transfer_EC", fontface = "bold", 
                    size = 5)


p2 <- DimPlot(sickle_cell, group.by = "Marker_Ann", label = F, pt.size = pt_size) + 
  theme(legend.position = "bottom") + ggtitle("Marker_Ann")
p2 <- LabelClusters(p2, id = "Marker_Ann", fontface = "bold", 
                    size = 5)


p3 <- DimPlot(sickle_cell, group.by = "Parse_Ann2", label = F, pt.size = pt_size) + 
  theme(legend.position = "bottom") + ggtitle("Parse_Ann2")
p3 <- LabelClusters(p3, id = "Parse_Ann2", fontface = "bold", 
                    size = 5)


png("UMAP_Annotation_transfer_EC.png", width = 17, height = 19, units = "in", res = 300) #7
print(p1)
dev.off()

png("UMAP_Marker_Ann.png", width = 17, height = 19, units = "in", res = 300) #7
print(p2)
dev.off()


png("UMAP_Parser_Ann2.png", width = 17, height = 19, units = "in", res = 300) #7
print(p3)
dev.off()


getwd()
```



```{r}
# refdata = Nature_RNA_reference$annot
#Stromal_parse_again$detailed_functional3
# Parsall_comb$general_group_nounk,




```




```{r}
#retrieved from 354 in thi .Rmd

sickle_cell$Parse_Ann2 <- NA

for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
    anchors <- FindTransferAnchors(
      reference = RNA_ScAnemia,         # Reference Seurat object (with known T cell labels)
  query = obj,                 # Query Seurat object (to which we want to transfer labels)
  normalization.method = "LogNormalize",       # Normalization method (use "SCT" if you did SCTransform or "SCT")
  reference.assay = "RNA",            # Assay in the reference Seurat object (usually "RNA" for raw or normalized counts)
  query.assay = "RNA",                # Assay in the query Seurat object
  dims = 1:30)
                #2- Second Tranfer labels
    #first label
    Bis_TRANSITIVE_Elh <- TransferData(anchorset = anchors,
                             refdata = RNA_ScAnemia$Parse_Ann2,
                             dims = 1:30
                             )
    obj$Parse_Ann2<- Bis_TRANSITIVE_Elh$predicted.id
    yes <- colnames(obj)
    sickle_cell$Parse_Ann2[yes] <- obj$Parse_Ann2[yes]
  
    assign(i, obj) 
}
```


#MARK ANN AGAIn


```{r}
for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
    obj <- get(i)
    anchors <- FindTransferAnchors(
      reference = Parsall_comb,         # Reference Seurat object (with known T cell labels)
  query = obj,                 # Query Seurat object (to which we want to transfer labels)
  normalization.method = "LogNormalize",       # Normalization method (use "SCT" if you did SCTransform or "SCT")
  reference.assay = "RNA",            # Assay in the reference Seurat object (usually "RNA" for raw or normalized counts)
  query.assay = "RNA",                # Assay in the query Seurat object
  dims = 1:30)
                #2- Second Tranfer labels

      Bis_TRANSITIVE_Elh <- TransferData(anchorset = anchors,
                             refdata = Parsall_comb$Marker_Ann,
                             dims = 1:30
                             )
    obj$Marker_Ann<- Bis_TRANSITIVE_Elh$predicted.id
    yes <- colnames(obj)
    sickle_cell$Marker_Ann[yes] <- obj$Marker_Ann[yes]
    assign(i, obj) 
}

```

# PARSE ANN2 of Vpcs

```{r}

#SEE GENERAL GROUP IN DATAFRAME-START
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
df1 <- as.data.frame(table(Vpc1$Parse_Ann2))
df2 <-  as.data.frame(table(Vpc2$Parse_Ann2))
df3 <- as.data.frame(table(Vpc3$Parse_Ann2))
df4 <- as.data.frame(table(Vpc4$Parse_Ann2))

rownames(df1) <- df1$Var1
rownames(df2) <- df2$Var1
rownames(df3) <- df3$Var1
rownames(df4) <- df4$Var1

#df$Var1 <- NULL
df1$Var1 <- NULL
df2$Var1 <- NULL
df3$Var1 <- NULL
df4$Var1 <- NULL

colnames(df1)[1] <- "WT Male"
colnames(df2)[1] <- "ScAnemia Male"
colnames(df3)[1] <- "WT Female"
colnames(df4)[1] <- "ScAnemia Female"


#ADDING AND CHANING POSITION OF ROWNAMES
df1 <- df1
df1 <- rbind(df1, 'Inflammatory Monocytes' = 0)

new_order <- c(rownames(df1)[1:10], "Inflammatory Monocytes", rownames(df1)[11:17])
# Reorder the DataFrame rows
df1 <- df1[new_order, , drop = FALSE]



df <- cbind(df1, df2, df3, df4)


library(ggplot2)
library(tidyr)

df$Cell_Type <- rownames(df)

# Convert data frame to long format
df_long <- pivot_longer(df, cols = c("WT Male", "ScAnemia Male", "WT Female", "ScAnemia Female"),
                        names_to = "Condition", values_to = "Count")


library(gridExtra)
library(ggplot2) 


# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = Count, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("ParseAnn2_sickle_number.png", width = 9, height = 11, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()


df_long$total <- NA
df_long$total[df_long$Condition == "WT Male"] <- 7973
df_long$total[df_long$Condition == "ScAnemia Male"] <- 7365
df_long$total[df_long$Condition == "WT Female"] <- 8886
df_long$total[df_long$Condition == "ScAnemia Female"] <- 7378


df_long$count_percent <- NA
df_long$count_percent <- (df_long$Count * 100) / df_long$total





p <- ggplot(df_long, aes(x = Cell_Type, y = count_percent, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f", count_percent)), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Percentage Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

png("Parse_Ann2_sickle_percentage.png", width = 9, height = 11, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()

```


#READ SICKLE CELL

```{r}


#Sickle Cell Anemia Project

#"C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/Sickle_Cell_Project_01_14_2025.Rmd"

sickle_cell <- readRDS("C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/sickle_cell_anemia_integrated_1_14_25.Rds")

sickle_list <- readRDS("C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/sickle_cell_list_anemia_integrated_1_14_25.Rds")

saveRDS(sickle_cell, "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/sickle_cell_anemia_integrated_1_14_25.Rds")

saveRDS(sickle_list, "C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/sickle_cell_list_anemia_integrated_1_14_25.Rds")


```





# CONDITION of Vpcs

```{r}


#SEE GENERAL GROUP IN DATAFRAME-START
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
#------------------------------------------------------------------------------------------------
df1 <- as.data.frame(table(Vpc1$Parse_Ann2))
df2 <-  as.data.frame(table(Vpc2$Parse_Ann2))
df3 <- as.data.frame(table(Vpc3$Parse_Ann2))
df4 <- as.data.frame(table(Vpc4$Parse_Ann2))

rownames(df1) <- df1$Var1
rownames(df2) <- df2$Var1
rownames(df3) <- df3$Var1
rownames(df4) <- df4$Var1

#df$Var1 <- NULL
df1$Var1 <- NULL
df2$Var1 <- NULL
df3$Var1 <- NULL
df4$Var1 <- NULL

colnames(df1)[1] <- "WT Male"
colnames(df2)[1] <- "ScAnemia Male"
colnames(df3)[1] <- "WT Female"
colnames(df4)[1] <- "ScAnemia Female"


#ADDING AND CHANING POSITION OF ROWNAMES
df1 <- df1
df1 <- rbind(df1, 'Inflammatory Monocytes' = 0)

new_order <- c(rownames(df1)[1:10], "Inflammatory Monocytes", rownames(df1)[11:17])
# Reorder the DataFrame rows
df1 <- df1[new_order, , drop = FALSE]



df <- cbind(df1, df2, df3, df4)


library(ggplot2)
library(tidyr)

df$Cell_Type <- rownames(df)

df$WT <- NA
df$ScAnemia <- NA
df$WT <- df$`WT Male` + df$`WT Female`
df$ScAnemia <- df$`ScAnemia Male` + df$`ScAnemia Female`



# Convert data frame to long format
df_long <- pivot_longer(df, cols = c("WT", "ScAnemia"),
                        names_to = "Condition", values_to = "Count")


library(gridExtra)
library(ggplot2) 


# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = Count, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("ParseAnn2_sickle_number_only_condition.png", width = 9, height = 11, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()

length(df$WT)
df_long$total <- NA
df_long$total[df_long$Condition == "WT"] <- 16859
df_long$total[df_long$Condition == "ScAnemia"] <- 14743



df_long$count_percent <- NA
df_long$count_percent <- (df_long$Count * 100) / df_long$total





p <- ggplot(df_long, aes(x = Cell_Type, y = count_percent, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f", count_percent)), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Percentage Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

png("ParseAnn2_sickle_percentage_only_condition.png", width = 9, height = 11, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()




```







```{r, fig.height=12, fig.width=15}
rownames(sickle_cell)[grep("Hba", rownames(sickle_cell))]

#[1] "Hbb-bt"= thalesemia, "Hbb-bs"= sickle_cell_anmia "Hbb-y"
#[1] "Hba-x"  "Hba-a2" "Hba-a1"
FeaturePlot(sickle_cell, "Hbb-bs", split.by = "condition")

DotPlot(sickle_cell,  "Hbb-bs", group.by = "condition", scale.min = 0)
DotPlot(sickle_cell,  "Hbb-bt", group.by = "condition", scale.min = 0)
DotPlot(sickle_cell,  "Hba-x", group.by = "condition", scale.min = 0)
DotPlot(sickle_cell,  "Hba-a1", group.by = "condition", scale.min = 0)
DotPlot(sickle_cell,  "Hba-a2", group.by = "condition", scale.min = 0)


DotPlot(RNA_ScAnemia,  "Hbb-bs", group.by = "group", scale.min = 0)
DotPlot(RNA_ScAnemia,  "Hbb-bt", group.by = "group", scale.min = 0)
DotPlot(RNA_ScAnemia,  "Hba-x", group.by = "group", scale.min = 0)
DotPlot(RNA_ScAnemia,  "Hba-a1", group.by = "group", scale.min = 0)
DotPlot(RNA_ScAnemia,  "Hba-a2", group.by = "group", scale.min = 0)


```







```{r}
sickle_cell$condition <- NA 

sickle_cell$condition <- ifelse(sickle_cell$group %in% c("WT Male", "WT Female"), 
                                "WT", 
                                "ScAnemia")
  getwd()
  
  table(sickle_cell$condition)
```

#yes
```{r}

NOS_flow_genes <- c(
  "Nos3",    # Endothelial nitric oxide synthase (eNOS)
  "Nos2",    # Inducible nitric oxide synthase (iNOS)
  "Vcam1",   # Vascular cell adhesion molecule 1
  "Gch1",    # GTP cyclohydrolase 1
  "Arg1",    # Arginase 1
  "Arg2",    # Arginase 2
  "Slc7a1",  # Cationic amino acid transporter 1
  "Edn1",    # Endothelin-1
  "Cyp2c9",  # Cytochrome P450 family 2 subfamily C member 9
  "Cyp2c19", # Cytochrome P450 family 2 subfamily C member 19
  "Akt1",    # Protein kinase B (Akt1)
  "Pde5a"    # Phosphodiesterase 5A
)


sca_lung_rat_genes <- c(
  # Hypoxia Response
  "Hif1a", "Epo", "Vegfa", 

  # Endothelial Dysfunction and Adhesion
  "Vcam1", "Icam1", "Sele", "Selplg",

  # Inflammation
  "Il6", "Tnf", "Nfkb1", "Cxcl8", "Ccl2", "Cxcl10",

  # Oxidative Stress
  "Nos3", "Sod1", "Gclc", 

  # Hemoglobin Regulation (Updated)
  "Hbb-bt", "Hbb-bs", "Hbs1l", "Hba-x", "Hba-a1", "Hba-a2", "Hbq1a", "Hbb-y", 

  # Lung-Enriched Genes
  "Sftpa1", "Sftpb", "Sftpc", "Ace", "Edn1"
)



DotPlot(sickle_cell, NOS_flow_genes, group.by = "condition", scale.min = 0) + coord_flip()
DotPlot(sickle_cell, NOS_flow_genes, group.by = "group", scale.min = 0) + coord_flip() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

yes <- c("WT Male", "ScAnemia Male", "WT Female", "ScAnemia Female")

yes1 <- unique(sickle_cell$group)
setdiff(yes, yes1)
sickle_cell$group <- factor(sickle_cell$group, yes)

#HEATMAP


```




```{r}

sca_lung_rat_genes <- c(
  # Hypoxia Response
  "Hif1a", "Epo", "Vegfa", 

  # Endothelial Dysfunction and Adhesion
  "Vcam1", "Icam1", "Sele", "Selplg",

  # Inflammation
  "Il6", "Tnf", "Nfkb1", "Cxcl8", "Ccl2", "Cxcl10",

  # Oxidative Stress
  "Nos3", "Sod1", "Gclc", 

  # Hemoglobin Regulation (Updated)
  "Hbb-bt", "Hbb-bs", "Hbs1l", "Hba-x", "Hba-a1", "Hba-a2", "Hbq1a", "Hbb-y", 

  # Lung-Enriched Genes
  "Sftpa1", "Sftpb", "Sftpc", "Ace", "Edn1"
)

#1-
# integrated
library(ComplexHeatmap)
library(dplyr)
library(Seurat)
expression_data_deg <- FetchData(sickle_cell, vars = sca_lung_rat_genes)
groups <- sickle_cell$group
#1: 
group_raw_averages <- compute_group_averages(expression_data_deg, groups)
#2:
#Scale -2 to +2
scaled_data_Deg <- apply(group_raw_averages[, -1], 2, scale_to_range)
# Convert the summarized data back to a matrix for heatmap
heatmap_matrix <- as.matrix(scaled_data_Deg)  # Exclude the group column (now rows are groups)
rownames(heatmap_matrix) <- group_raw_averages$group  # Set group names as row names
heatmap_matrix <- t(heatmap_matrix)
# Generate the heatmap
p <- Heatmap(
  heatmap_matrix,
  name = "Expression",
  cluster_rows = FALSE,         # Cluster genes
  cluster_columns = FALSE,      # Cluster groups
  show_row_names = TRUE,       # Show gene names
  show_column_names = TRUE,    # Show group names
  column_names_side = "top",
  row_names_side = "left",
   row_names_gp = gpar(fontsize = 10)  # General text size for row names
    # Only label rows in 'markers'
)
print(p)

```

# THE GROUPS ADDED!!!!

```{r, fig.height=10, fig.width=6}
# Define gene categories
gene_categories <- c(
  # Hypoxia Response
  "Hif1a" = "Hypoxia Response", "Epo" = "Hypoxia Response", "Vegfa" = "Hypoxia Response",
  
  # Endothelial Dysfunction and Adhesion
  "Vcam1" = "Endothelial Dysfunction and Adhesion", 
  "Icam1" = "Endothelial Dysfunction and Adhesion",
  "Sele" = "Endothelial Dysfunction and Adhesion", 
  "Selplg" = "Endothelial Dysfunction and Adhesion",
  
  # Inflammation
  "Il6" = "Inflammation", "Tnf" = "Inflammation", "Nfkb1" = "Inflammation", 
  "Cxcl8" = "Inflammation", "Ccl2" = "Inflammation", "Cxcl10" = "Inflammation",
  
  # Oxidative Stress
  "Nos3" = "Oxidative Stress", "Sod1" = "Oxidative Stress", "Gclc" = "Oxidative Stress",
  
  # Hemoglobin Regulation
  "Hbb-bt" = "Hemoglobin Regulation", "Hbb-bs" = "Hemoglobin Regulation",
  "Hbs1l" = "Hemoglobin Regulation", "Hba-x" = "Hemoglobin Regulation",
  "Hba-a1" = "Hemoglobin Regulation", "Hba-a2" = "Hemoglobin Regulation",
  "Hbq1a" = "Hemoglobin Regulation", "Hbb-y" = "Hemoglobin Regulation",
  
  # Lung-Enriched Genes
  "Sftpa1" = "Lung-Enriched Genes", "Sftpb" = "Lung-Enriched Genes",
  "Sftpc" = "Lung-Enriched Genes", "Ace" = "Lung-Enriched Genes", 
  "Edn1" = "Lung-Enriched Genes"
)

# Create row annotations
row_annotation <- rowAnnotation(
  Category = factor(gene_categories[rownames(heatmap_matrix)]),
  col = list(Category = c(
    "Hypoxia Response" = "blue",
    "Endothelial Dysfunction and Adhesion" = "green",
    "Inflammation" = "red",
    "Oxidative Stress" = "purple",
    "Hemoglobin Regulation" = "orange",
    "Lung-Enriched Genes" = "brown"
  ))
)

# Generate the heatmap
p <- Heatmap(
  heatmap_matrix,
  name = "Expression",
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  column_names_side = "top",
  row_names_side = "left",
  row_names_gp = gpar(fontsize = 10),
  left_annotation = row_annotation
)

# Print the heatmap
print(p)

```

#LASTCALL

```{r, fig.width=17, fig.height=19}
library(ggplot2)
p3 <- DimPlot(sickle_cell, group.by = "Last_Call", label = F) + 
  theme(legend.position = "bottom") + ggtitle("Last_Call")
LabelClusters(p3, id = "Last_Call", fontface = "bold", 
                    size = 5)
```

```{r, fig.width=30, fig.height=12}
library(ggplot2)
DimPlot(sickle_cell, group.by = "Last_Call", label = F, split.by = "group", raster = F) + 
  theme(legend.position = "bottom") + ggtitle("Last_Call") 

```

```{r}
FeaturePlot(sickle_cell, "Gja5", order = F)
```
```{r, fig.width=17, fig.height=19}
library(ggplot2)
p3 <- DimPlot(sickle_cell, group.by = "origin_of_cells_nounk", label = F) + 
  theme(legend.position = "bottom") + ggtitle("origin_of_cells_nounk")
LabelClusters(p3, id = "origin_of_cells_nounk", fontface = "bold", 
                    size = 5)

library(ggplot2)
p3 <- DimPlot(sickle_cell, group.by = "origin_of_cells_with_unk", label = F) + 
  theme(legend.position = "bottom") + ggtitle("origin_of_cells_with_unk")
LabelClusters(p3, id = "origin_of_cells_with_unk", fontface = "bold", 
                    size = 5)

FeaturePlot(sickle_cell, "Pecam1", order = F)

```
```{r,fig.height=7, fig.width=15}
DimPlot(sickle_cell, group.by = "Last_Call", label = F, cells.highlight = colnames(sickle_cell)[sickle_cell$Last_Call == "Alveolar Macrophages"], split.by = "group") + 
  theme(legend.position = "bottom") + ggtitle("Alveolar Macrophages in Last_Call")

DimPlot(sickle_cell, group.by = "Parse_Ann2", label = F, cells.highlight = colnames(sickle_cell)[sickle_cell$Parse_Ann2 == "T cells"], split.by = "group") + 
  theme(legend.position = "bottom") + ggtitle("T cells in Last_Call")

DimPlot(sickle_cell, group.by = "Parse_Ann2", label = F, cells.highlight = colnames(sickle_cell)[sickle_cell$Parse_Ann2 == "Pericytes"], split.by = "group") + 
  theme(legend.position = "bottom") + ggtitle("Pericytes in Last_Call")

DimPlot(sickle_cell, group.by = "Last_Call", label = F, cells.highlight = colnames(sickle_cell)[sickle_cell$Last_Call == "Myofibroblasts"], split.by = "group") + 
  theme(legend.position = "bottom") + ggtitle("Myofibroblasts in Last_Call")


table(sickle_cell$Parse_Ann2)
```



```{r, fig.height=7, fig.width=25}
# Hemoglobin Regulation
 Hbb_genes <- c("Hbb-bt", "Hbb-bs",
  "Hbs1l", "Hba-x",
  "Hba-a1", "Hba-a2",
  "Hbq1a", "Hbb-y")

FeaturePlot(sickle_cell,Hbb_genes, #split.by = "group"
            )

FeaturePlot(sickle_cell,"Hba-a2", split.by = "group"
            )


DimPlot(sickle_cell, label = T, label.size = 10) 

Hbb <- subset(sickle_cell, seurat_clusters == 10)

DimPlot(Hbb, label = T, label.size = 10) 
FeaturePlot(Hbb,Hbb_genes)


markersc <- c("Cd68", "Cd163", "Mrc1", "Marco", "Trem2", "Lamp1", "Lamp2", "Map1lc3b", "Vsig4")
FeaturePlot(sickle_cell,markersc)

FeaturePlot(sickle_cell,"Map1lc3b", split.by = "group")

LC3B (MAP1LC3B)	Autophagy-related marker (often upregulated in macrophages digesting engulfed material)

#kaldiikkkkkkkkkkkkkkkkkkkkk

```





```{r}

library(ggplot2)
DotPlot(Hbb, Hbb_genes, group.by = "group") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
DotPlot(Hbb, c(
   "Hbb-y"), group.by = "group") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```



# EC_flex ARTICLE REVISION MARKERS markers

```{r}
DimPlot(EC_flex_3, group.by = "Annotation")
#EndIT EC
DefaultAssay(EC_flex_3) <- "SCT"
FeaturePlot(EC_flex_3, "TRBC2")
FeaturePlot(EC_flex_3, "CD2")
FeaturePlot(EC_flex_3, "ICOS", order = T)
FeaturePlot(EC_flex_3, "CRTAM")
FeaturePlot(EC_flex_3, "CD6")


#Infla EC
FeaturePlot(EC_flex_3, "SELE")
FeaturePlot(EC_flex_3, "IL6")

FeaturePlot(EC_flex_3, "ACKR1")

#prolif EC
FeaturePlot(EC_flex_3, "SEMA3G")
FeaturePlot(EC_flex_3, "GJA5")

#EC3
FeaturePlot(EC_flex_3, "RGCC")
FeaturePlot(EC_flex_3, "CD36")
FeaturePlot(EC_flex_3, "RBP7")
FeaturePlot(EC_flex_3, "CD300LG")
FeaturePlot(EC_flex_3, "CA4")
FeaturePlot(EC_flex_3, "TNFRSF4")
FeaturePlot(EC_flex_3, "MT1M")


FeaturePlot(EC_flex_3, "FABP4")


```



# NULLLLLLLL LETS CHECK SEX DIFFERENCE

```{r}
sickle_cell$ <- factor(c(rep("VPC1", n1), rep("VPC2", n2), rep("VPC3", n3), rep("VPC4", n4)))


rm(list = setdiff(ls(), "sickle_cell"))
```


#----------------------------------------------------------------------------------------------------#----------------------------------------------------------------------------------------------------#----------------------------------------------------------------------------------------------------#--------------------------------------------------------------------------------------------------------------------------
```{r}
erythroid_markers <- c("Tfrc",  # Transferrin receptor (iron uptake, erythroid precursor marker)
                       "Kit",  # Stem cell factor receptor (early erythroid marker)
                       "Klf1",  # Key erythroid transcription factor
                       "Gata1",  # Master regulator of erythroid lineage
                       "Hba-a2", "Hbb",  # Hemoglobin subunits (definitive erythroid markers)
                       "Gypa",  # Glycophorin A (mature RBC marker)
                       "Fech",  # Ferrochelatase (heme biosynthesis)
                       "Ahsp",  # Stabilizes hemoglobin
                       "Slc2a39", "Slc1a4",  # Iron and amino acid transporters for erythropoiesis
                       "Gpx1",  # Protects erythroid cells from oxidative stress
                       "Bnip3l", "Map1lc3b")  # Autophagy genes for RBC maturation
macrophage_markers <- c("Cd68",  # General macrophage marker
                        "Cd163",  # Scavenger receptor (high in RBC-clearing macrophages)
                        "Marco",  # Macrophage receptor involved in phagocytosis
                        "Mrc1",  # CD206, mannose receptor (phagocytic macrophages)
                        "Fcer1g",  # Fc receptor gamma (involved in phagocytosis)
                        "Adgre1",  # F4/80, classic macrophage marker
                        "Lyz2",  # Lysozyme, expressed in macrophages
                        "Hamp",  # Hepcidin, involved in iron metabolism
                        "Hmox1",  # Heme oxygenase 1, involved in RBC breakdown
                        "Slc40a1",  # Ferroportin, regulates iron export
                        "Fcgr1", "Fcgr2b")  # Fc receptors for immune functions

```


#

```{r}

p <- DimPlot(sickle_cell, label = F, group.by = "Parse_Ann") + NoLegend()
    LabelClusters(p, id = "Parse_Ann", fontface = "bold", size = 5)
    
meta_data <- sickle_cell@meta.data


```



    ```{r}

# ANALYSIS OF SICKLE CELL ABOVE 2-17-25




```

#CHECK

```{r}

#Idents(Endo_ScAnemia) <- Endo_ScAnemia$group
#Endo_ScAnemia <- JoinLayers(Endo_ScAnemia, assay = "RNA")
find_and_save_markers <- function(seurat_obj) {
  # Define the identifier pairs within the function
  for (i in sub_endo) {
    seurat_obj <- subset(Endo_ScAnemia, annot1 %in% i)
      ident_pairs <- list(
    c("WT Female", "WT_Male"),
    c("ScAnemia Female", "WT Female"),
    c("ScAnemia Male", "WT_Male"),
    c("ScAnemia Female", "ScAnemia Male")
      )
        # Loop through each pair and find markers
        for (pair in ident_pairs) {
              # Extract ident.1 and ident.2 from the pair
              ident_1 <- pair[1]
                  ident_2 <- pair[2]
                      # Find markers
                      markers <- FindMarkers(seurat_obj, ident.1 = ident_1, ident.2 = ident_2, only.pos = FALSE)
                      
                          # Select top 33 markers based on avg_log2FC
                          top_markers <- markers %>% 
                                  filter(p_val_adj < 0.05) %>% 
                                  slice_max(order_by = avg_log2FC, n = 50)
                          
                              # Create a filename based on the identifiers
                              filename <- paste0(i, "_", ident_1, "_vs_", ident_2, ".csv")
                                  # Write the markers to a CSV file
                                  write.csv(top_markers, filename)
                                      # Print a message indicating progress
                                      print(paste("Markers for", ident_1, "vs", ident_2, "saved to", filename))
        }
  }
}

# Run the function
find_and_save_markers(Endo_ScAnemia)



```

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)  # For human, use org.Mm.eg.db for mouse
library(enrichplot)
library(ggplot2)

#1558 pATHWAY ANALYSIS
getwd()

```


#RASA3 - Prohaska, Clare Colette

```{r, fig.height=4, fig.width=7}
library(Seurat)
DimPlot(sickle_cell)

DotPlot(sickle_cell, "Rasa3")

cell_pops <- unique(sickle_cell$Parse_Ann2)
for(i in cell_pops){
  Idents(sickle_cell) <- sickle_cell$Parse_Ann2
p <- DotPlot(sickle_cell, idents = i, "Rasa3", split.by = "condition", scale.min = 0)
png(paste0("./RASA3/", i, "_RASA3.png"), width = 7, height = 5, res = 300, units = "in" )
print(p)
dev.off()
}



cell_pops <- unique(sickle_cell$Parse_Ann2)
for (i in cell_pops) {
  Idents(sickle_cell) <- sickle_cell$Parse_Condition
  
  # Select cells that contain 'i' in their names across different groups
  selected_cells <- grep(i, sickle_cell$Parse_Condition, value = TRUE) 

  # If there are matching cells, generate a DotPlot
  if (length(selected_cells) > 0) {
    p <- DotPlot(sickle_cell, features = "Rasa3", idents = selected_cells, scale.min = 0)

    # Save the plot
    png(paste0("./RASA3_1/", i, "_RASA3.png"), width = 7, height = 5, res = 300, units = "in")
    print(p)
    dev.off()
  }
}

sickle_cell$Parse_Condition <- NA
sickle_cell$Parse_Condition <- paste(sickle_cell$condition, sickle_cell$Parse_Ann2)


getwd()
LinkedDimPlot(WSA_LngSP10193346)
LinkedFeaturePlot(anchored_spas, "ACKR1"
                  , alpha = 0.1
                  )

LinkedDimPlot(anchored_spas)
print(p)

LinkedFeaturePlot(anchored_spas, "ACKR1")

SpatialDimPlot(WSA_LngSP10193346, alpha = 1)
library(ggplot2)
SpatialFeaturePlot(WSA_LngSP10193346, "PECAM1") + theme(legend.text = element_text(size = 5))

rm(anchored_spas)
?LinkedDimPlot

```


SpatialFeaturePlot(WSA_LngSP10193346, "PECAM1", pt.size = 4)
SpatialFeaturePlot(WSA_LngSP10193346, "PECAM1", pt.size = 5) + theme(legend.text = element_text(size = 5))

SpatialDimPlot(WSA_LngSP10193346, pt.size = 5, label = T) + theme(legend.text = element_text(size = 5)) + NoLegend()

?SpatialDimPlot



#DOIT



#WORKED
```{r}


library(Seurat)
library(ggplot2)

# List of identity pairs for DEG comparisons
ident_pairs <- list(
  c("WT Female", "WT Male"),
  c("ScAnemia Female", "WT Female"),
  c("ScAnemia Male", "WT Male"),
  c("ScAnemia Female", "ScAnemia Male"),
  c("WT Male", "WT Female"),
  c("WT Female", "ScAnemia Female"),
  c("WT Male", "ScAnemia Male"),
  c("ScAnemia Male", "ScAnemia Female")
)

# Create a dataframe to store results
deg_counts_df <- data.frame(Cell_Group = character(),
                            Pair_Comparison = character(),
                            DEG_Count = numeric(),
                            stringsAsFactors = FALSE)

# Loop through unique cell populations
cell_pops <- unique(sickle_cell$Parse_Ann2)

for (i in cell_pops) {
  # Subset the dataset for the given cell group
  sickle_subset <- subset(sickle_cell, Parse_Ann2 == i)
  
  # Loop through each identity pair
  for (pair in ident_pairs) {
    ident_1 <- pair[1]
    ident_2 <- pair[2]
    
    # Check if both identities exist in the subset
    if (all(c(ident_1, ident_2) %in% sickle_subset$group)) {
      
      # Perform DEG analysis
      degs <- FindMarkers(sickle_subset, ident.1 = ident_2, ident.2 = ident_1, 
                          assay = "RNA", group.by = "group", only.pos = TRUE)
      
      # Filter significant DEGs (adjusted p-value < 0.05)
      significant_degs <- degs[degs$p_val_adj < 0.05, ]
      deg_count <- nrow(significant_degs)  # Count significant DEGs
      
      # Store results in dataframe
      deg_counts_df <- rbind(deg_counts_df, 
                             data.frame(Cell_Group = i, 
                                        Pair_Comparison = paste(ident_1, "vs", ident_2), 
                                        DEG_Count = deg_count))
    }
  }
}

# Save DEG counts to a CSV file
write.csv(deg_counts_df, "DEG_counts_per_cell_group.csv", row.names = FALSE)

# Plot the results
ggplot(deg_counts_df, aes(x = Pair_Comparison, y = DEG_Count, fill = Cell_Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  theme_minimal() +
  labs(title = "DEG Counts per Cell Group & Pair Comparison", 
       x = "Pair Comparison", y = "Number of DEGs") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

getwd()




```



```{r}
library(dplyr)
library(tidyr)

deg_counts_df <- read.csv("DEG_counts_per_cell_group.csv")
# Split the "Pair_Comparison" column into two new columns ("From" and "To")
deg_counts_df <- deg_counts_df %>%
  separate(Pair_Comparison, into = c("To", "From"), sep = " vs ", remove = FALSE) %>%
  select(Cell_Group, To, From, Pair_Comparison, DEG_Count)  # Reorder columns

df <- deg_counts_df
df <- df[, -4]
df <- df %>%
  select(1, 3, 2, everything())
df_conserve <- df
rm(d)
```

#WORKING making graph- ELHAMDULILLAH
```{r}
 # Filter for one cell type
  df <- df_conserve[df_conserve$Cell_Group == "AT2 cells", ]

# Assign x, y coordinates for plotting
positions <- data.frame(
  Group = unique(c(df$From, df$To)),
  x = c(0, 2, -2, 0),
  y = c(2, 0, 0, -2)
)

# Merge coordinates to main data
df <- df %>%
  left_join(positions, by = c("From" = "Group")) %>%
  rename(x_start = x, y_start = y) %>%
  left_join(positions, by = c("To" = "Group")) %>%
  rename(x_end = x, y_end = y)


df$x_end <- df$x_start + (df$x_end - df$x_start) * 0.75  
df$y_end <- df$y_start + (df$y_end - df$y_start) * 0.75  

# Plot the directed graph with arrows
ggplot() +
  # Arrows for connections
  geom_curve(data = df, aes(x = x_start, y = y_start, xend = x_end, yend = y_end), 
             curvature = -0.5, arrow = arrow(length = unit(0.2, "inches")), size = 1.2, color = "red") +

  # Labels for DEG counts above the arrows
  geom_text(data = df, aes(x = (x_start + x_end) / 2, y = (y_start + y_end) / 2 + ifelse(y_start > y_end, 0.5, -0.5), 
                                      label = paste(DEG_Count, "degs")), 
            color = "black", fontface = "bold", size = 5) +

  # Labels for each group (WT Male, WT Female, etc.)
  geom_label(data = positions, aes(x = x, y = y, label = Group), 
             color = "black", fontface = "bold", fill = "yellow", size = 5) +

  
  # EXPAND GRAPH AREA
  xlim(min(df$x_start) - 1, max(df$x_end) + 1) +  # Increase left/right range
  ylim(min(df$y_start) - 1, max(df$y_end) + 1) +  # Increase top/bottom range
  # Customize theme
  theme_void() +
  labs(title = "DEG Count Flowchart", x = "", y = "")




```


```{r, fig.height=8, fig.width=8}

library(ggplot2)
library(dplyr)

# Ensure directory exists
dir.create("./DEGs", showWarnings = FALSE)

cell_pop <- unique(df_conserve$Cell_Group)

for (i in cell_pop) {
  # Filter for one cell type
  df_subset <- df_conserve[df_conserve$Cell_Group == i, ]

  # Get unique groups from 'From' and 'To' columns
  unique_groups <- unique(c(df_subset$From, df_subset$To))
  n <- length(unique_groups)  # Number of unique groups

  # Dynamically assign positions (circular layout)
  angles <- seq(0, 2 * pi, length.out = n + 1)[-1]  # Evenly spaced angles
  positions <- data.frame(
    Group = unique_groups,
    x = cos(angles) * 3,  # Spread in a circular shape
    y = sin(angles) * 3
  )

  # Merge coordinates to main data
  df_subset <- df_subset %>%
    left_join(positions, by = c("From" = "Group")) %>%
    rename(x_start = x, y_start = y) %>%
    left_join(positions, by = c("To" = "Group")) %>%
    rename(x_end = x, y_end = y)

  # Shorten arrows to avoid overlap
  df_subset$x_end <- df_subset$x_start + (df_subset$x_end - df_subset$x_start) * 0.75  
  df_subset$y_end <- df_subset$y_start + (df_subset$y_end - df_subset$y_start) * 0.75  

  # Define axis limits safely
  x_limits <- range(df_subset$x_start, df_subset$x_end) + c(-1, 1)
  y_limits <- range(df_subset$y_start, df_subset$y_end) + c(-1, 1)

  # Create the plot
  p <- ggplot() +
    # Arrows for connections
    geom_curve(data = df_subset, aes(x = x_start, y = y_start, xend = x_end, yend = y_end), 
               curvature = -0.4, arrow = arrow(length = unit(0.2, "inches")), size = 1.2, color = "red") +

    # Labels for DEG counts
    geom_text(data = df_subset, aes(
        x = (x_start + x_end) / 2 + ifelse(x_start > x_end, 0.3, -0.3),  
        y = (y_start + y_end) / 2 + ifelse(y_start > y_end, 0.5, -0.5),  
        label = paste(DEG_Count, "degs")), 
        color = "black", fontface = "bold", size = 5) +

    # Labels for each group
    geom_label(data = positions, aes(x = x, y = y, label = Group), 
               color = "black", fontface = "bold", fill = "yellow", size = 5) +

    # Expand graph area
    xlim(x_limits) + ylim(y_limits) +

    # Customize theme
    theme_void() +
    labs(title = paste("DEG Count Flowchart for", i), x = "", y = "")

  # Save the plot
  png(paste0("./DEGs/", i, ".png"), height = 8, width = 8, res = 300, units = "in")
  print(p)
  dev.off()
}


write.csv(df_conserve, "all_degs_per_cell_condition.csv")

```


#NETWORK WORKED
```{r}
library(igraph)
df_split <- split(df_conserve, df_conserve$Cell_Group)

for (cell_type in names(df_split)) {
  df_sub <- df_split[[cell_type]]
  graph <- graph_from_data_frame(df_sub[, c("From", "To")], directed = TRUE)
  png(paste0("./Network_", cell_type, ".png"), width = 800, height = 800, res = 150)
  plot(graph, vertex.size = 15, edge.width = df_sub$DEG_Count / max(df_sub$DEG_Count) * 5)
  dev.off()
}

```

#CHORD WORKED
```{r}
library(circlize)
df_split <- split(df_conserve, df_conserve$Cell_Group)  # Separate by Cell_Group

for (cell_type in names(df_split)) {
  df_sub <- df_split[[cell_type]]
  png(paste0("./Chord_", cell_type, ".png"), width = 800, height = 800, res = 150)
  chordDiagram(df_sub[, c("From", "To", "DEG_Count")], 
               transparency = 0.5, grid.col = c("WT" = "blue", "Sickle" = "red"))
  dev.off()
}
```



# ANALYSIS OF SICKLE CELL ABOVE


#----------------------------------------------------------------------------------------------------#----------------------------------------------------------------------------------------------------#----------------------------------------------------------------------------------------------------#--------------------------------------------------------------------------------------------------------------------------



































































#Smc_pericytes

```{r, fig.height=8, fig.width=15}

library(Seurat)
sickle_cell <- JoinLayers(sickle_cell, assay = "RNA")

sickle_cell_ 


Idents(sickle_cell) <- sickle_cell$group


Diet_sickle <- DietSeurat(sickle_cell, scale.data = )

Diet_sickle@graphs$integrated_nn <- NULL
Diet_sickle@assays[["RNA"]]@layers[["scale.data.4"]] <- NULL
Diet_sickle@assays[["RNA"]]@meta.data <- NULL

rm(Diet_sickle)
```




```{r}


sickle_cell$group <- factor(sickle_cell$group, levels = c("WT Male", "ScAnemia Male", "WT Female", "ScAnemia Female"))

```

```{r}

library(dplyr)
      ident_pairs <- list(
    c("ScAnemia Male", "WT Male"),
    c("ScAnemia Female", "ScAnemia Male"),
    c("WT Female", "WT Male"),
    c("ScAnemia Female", "WT Female"),
    c("WT Male", "ScAnemia Female"),
    c("ScAnemia Female", "WT Male")
      )
        # Loop through each pair and find markers
        for (pair in ident_pairs) {
              # Extract ident.1 and ident.2 from the pair
              ident_1 <- pair[1]
                  ident_2 <- pair[2]
                      # Find markers
                      markers <- FindMarkers(sickle_cell, ident.1 = ident_1, ident.2 = ident_2, only.pos = FALSE)
                      
                          # Select top 33 markers based on avg_log2FC
                          top_markers <- markers %>% 
                                  filter(p_val_adj < 0.05) 
                          #%>% 
                                  #slice_max(order_by = avg_log2FC
                                            #, n = 50
                                            #)
                          
                              # Create a filename based on the identifiers
                              filename <- paste0("Whole_lung", "_", ident_1, "_vs_", ident_2, ".csv")
                                  # Write the markers to a CSV file
                                  write.csv(top_markers, filename)
                                      # Print a message indicating progress
                                      print(paste("Markers for", ident_1, "vs", ident_2, "saved to", filename))
        }

# Run the function
#find_and_save_markers(Endo_ScAnemia)



```


```{r}
for (i in 1:4) {
  
 p <-  DotPlot(sickle_list[[i]], Hbb_genes, group.by = "orig.ident") + theme(axis.text.x = element_text(angle = 45, hjust =1))
 print(p)
  
}

 DotPlot(sickle_list[[3]], Hbb_genes, group.by = "orig.ident") + theme(axis.text.x = element_text(angle = 45, hjust =1))
 
 DotPlot(sickle_cell, Hbb_genes, group.by = "orig.ident") + theme(axis.text.x = element_text(angle = 45, hjust =1))
  DotPlot(sickle_cell, Hbb_genes, group.by = "group") + theme(axis.text.x = element_text(angle = 45, hjust =1))

```

#FindAllMarkers


```{r, fig.height=25, fig.width=19}

sickle_cell_markers <- FindAllMarkers(sickle_cell, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 1, return.thresh = 0.5, group.by = "group")

DoHeatmap(sickle_cell, sickle_cell_markers$gene, group.by = "group")



```

```{r}

#1-
# integrated
library(ComplexHeatmap)
library(dplyr)
library(Seurat)
expression_data_deg <- FetchData(sickle_cell, vars = sickle_cell_markers$gene)
groups <- sickle_cell$group
#1: 
group_raw_averages <- compute_group_averages(expression_data_deg, groups)
#2:
#Scale -2 to +2
scaled_data_Deg <- apply(group_raw_averages[, -1], 2, scale_to_range)
# Convert the summarized data back to a matrix for heatmap
heatmap_matrix <- as.matrix(scaled_data_Deg)  # Exclude the group column (now rows are groups)
rownames(heatmap_matrix) <- group_raw_averages$group  # Set group names as row names
heatmap_matrix <- t(heatmap_matrix)
# Generate the heatmap
p <- Heatmap(
  heatmap_matrix,
  name = "Expression",
  cluster_rows = FALSE,         # Cluster genes
  cluster_columns = FALSE,      # Cluster groups
  show_row_names = TRUE,       # Show gene names
  show_column_names = TRUE,    # Show group names
  column_names_side = "top",
  row_names_side = "left",
   row_names_gp = gpar(fontsize = 10)  # General text size for row names
    # Only label rows in 'markers'
)
print(p)

```













```{r}

Idents(smc) <- smc$group
#Endo_ScAnemia <- JoinLayers(Endo_ScAnemia, assay = "RNA")
find_and_save_markers <- function(seurat_obj) {
  # Define the identifier pairs within the function
      ident_pairs <- list(
    c("WT Female", "WT_Male"),
    c("ScAnemia Female", "WT Female"),
    c("ScAnemia Male", "WT_Male"),
    c("ScAnemia Female", "ScAnemia Male")
      )
        # Loop through each pair and find markers
        for (pair in ident_pairs) {
              # Extract ident.1 and ident.2 from the pair
              ident_1 <- pair[1]
                  ident_2 <- pair[2]
                      # Find markers
                      markers <- FindMarkers(seurat_obj, ident.1 = ident_1, ident.2 = ident_2, only.pos = FALSE)
                      
                          # Select top 33 markers based on avg_log2FC
                          top_markers <- markers %>% 
                                  filter(p_val_adj < 0.05) %>% 
                                  #filter(p_val_adj < 0.05) %>% 
                                  #slice_max(order_by = avg_log2FC, n = 50)
                                  slice_max(order_by = p_val_adj, n = 20)
                          
                              # Create a filename based on the identifiers
                              filename <- paste0("SMC_ns_", "_", ident_1, "_vs_", ident_2, ".csv")
                                  # Write the markers to a CSV file
                                  write.csv(top_markers, filename)
                                      # Print a message indicating progress
                                      print(paste("Markers for", ident_1, "vs", ident_2, "saved to", filename))
        }
  }


# Run the function
find_and_save_markers(smc)

```



```{r, fig.height=6, fig.width=19}
library(clusterProfiler)
library(org.Hs.eg.db)  # For human, use org.Mm.eg.db for mouse
library(enrichplot)
library(ggplot2)

DimPlot(Endo_ScAnemia, group.by = "annot1", label = T, split.by = "group")

#Constructing the group name with ifelse
Endo_ScAnemia$group <- ifelse(grepl("_1$", colnames(Endo_ScAnemia)), "WT Normoxia",
                               ifelse(grepl("_2$", colnames(Endo_ScAnemia)), "WT Hypoxia",
                                      ifelse(grepl("_3$", colnames(Endo_ScAnemia)), "ScAnemia Normoxia", "ScAnemia Hypoxia")))
```
```{r, fig.height=19, fig.width=19}


p <- DimPlot(sickle_cell, label = F, group.by = "Parse_Ann") + NoLegend()
    LabelClusters(p, id = "Parse_Ann", fontface = "bold", size = 5)
    
    table(smc$group)

```


```{r, fig.height=7, fig.width=8}

FeaturePlot(sickle_cell, "ScAnemia", order = F)
FeaturePlot(ScAnemia_perismc, "ScAnemia", order = T, split.by = "group")
FeaturePlot(Stromal_parse_again, "ScAnemia", order = T)
DimPlot(Stromal_parse_again, raster = F, group.by = "detailed_functional3", label = T)

DotPlot(sickle_cell, "ScAnemia", group.by = "group", split.by = "Parse_general_group2", cols = c("red", "blue", "green", "orange", "yellow"))

AT1 <- subset(sickle_cell, Parse_Ann %in% "AT1 cells")
DimPlot(ScAnemia_perismc)


DotPlot(AT1, "ScAnemia", group.by = "group")

DotPlot(ScAnemia_perismc, "ScAnemia", group.by = "group")


smc_tr <- subset(ScAnemia_perismc, Parse_StromAnn2 %in% c("Smooth muscle", "Transitional Pericytes"))


DotPlot(ScAnemia_perismc, "Sphk2", group.by = "group")

peri <- subset(ScAnemia_perismc, Parse_StromAnn2 %in% c("Pericytes"))

table(ScAnemia_perismc$Parse_StromAnn2)



DotPlot(Endo_ScAnemia, "Sphk2", group.by = "group")


rownames(sickle_cell)[grep("Sk", rownames(sickle_cell))]


for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
  obj <- get(i)
  p <- DotPlot(obj, "ScAnemia")
  print(p)
}

DimPlot(Vpc1, label = T)


for (i in list("Vpc1", "Vpc2", "Vpc3", "Vpc4")){
  obj <- get(i)
  p1 <- DimPlot(Vpc1, label = T)
  p2 <- DimPlot(Vpc1, label = T, group.by = "Parse_Ann_i") + NoLegend()
  print(p1)
  print(p2)
}



DimPlot(Vpc1, label = T, group.by = "Parse_Ann_i") + NoLegend()

rm(list = ls())

DefaultAssay(Integ_SCT_mouse) <- "SCT"
FeaturePlot(Integ_SCT_mouse, "ScAnemia")

DimPlot(Integ_SCT_mouse, group.by = "IntegAnn", label = T) + NoLegend()


getwd()

```


#11/20/2024 DUTIES

```{r}

genes <- c("ScAnemia", "Sphk2", "Spp1", "Spp2", "Cers1", "Cers2", "Cers3", "Cers4", "Cers5", "Cers6", "Degs1", "Ugcg","Sptlc1", "Sgpl1", "S1pr1", "S1pr2", "S1pr3", "S1pr4", "S1pr5", "Igsf3", "Smpd1", "Galc", "Gba", "Smpd2", "Smpd3",  "Asah1", "Asah2",  "Acer1", "Acer2", "Acer3",  "Sgms1", "Sgms2","Ifng", "Cxcl10", "Col4a3bp", "Lpp")
#Whole lung
  #featureplot
for (i in genes){
  f <- paste0("featureplots/Whole_lung_", i, ".png")
  png(f, width = 19, height = 5, units = "in", res = 300)
  p <- FeaturePlot(sickle_cell, features = i, split.by = "group")
  print(p)
  dev.off()
}
  
  #dotplot
for (i in genes){
  f <- paste0("dotplots/Whole_lung_", i, ".png")
  png(f, width = 7, height = 5, units = "in", res = 300)
  p <- DotPlot(sickle_cell, features = i, group.by = "group1")
  print(p)
  dev.off()
}


#Smooth muscle-pericytes
  #featureplot
for (i in genes){
  f <- paste0("featureplots/Smc_pericyte_", i, ".png")
  png(f, width = 15, height = 5, units = "in", res = 300)
  p <- FeaturePlot(ScAnemia_perismc, features = i, split.by = "group")
  print(p)
  dev.off()
}
  
  #dotplot
for (i in genes){
  f <- paste0("dotplots/Smc_pericyte_", i, ".png")
  png(f, width = 7, height = 5, units = "in", res = 300)
  p <- DotPlot(ScAnemia_perismc, features = i, group.by = "group1")
  print(p)
  dev.off()
}



#Only Smooth muscle-
  #featureplot
for (i in genes){
  f <- paste0("featureplots/Only_Smc_", i, ".png")
  png(f, width = 15, height = 5, units = "in", res = 300)
  p <- FeaturePlot(smc, features = i, split.by = "group")
  print(p)
  dev.off()
}
  
  #dotplot
for (i in genes){
  f <- paste0("dotplots/Only_Smc_", i, ".png")
  png(f, width = 7, height = 5, units = "in", res = 300)
  p <- DotPlot(smc, features = i, group.by = "group1")
  print(p)
  dev.off()
}
  

#Smc_w_transitional
  #featureplot
for (i in genes){
  f <- paste0("featureplots/Smc_w_transitional_", i, ".png")
  png(f, width = 15, height = 5, units = "in", res = 300)
  p <- FeaturePlot(smc, features = i, split.by = "group")
  print(p)
  dev.off()
}
  
  #dotplot
for (i in genes){
  f <- paste0("dotplots/Smc_w_transitional_", i, ".png")
  png(f, width = 7, height = 5, units = "in", res = 300)
  p <- DotPlot(smc, features = i, group.by = "group1")
  print(p)
  dev.off()
}
  

#Endothelial cells


for (i in genes){
  f <- paste0("featureplots/Endothelial_subset_featureplots", i, ".png")
  png(f, width = 19, height = 5, units = "in", res = 300)
  p <- tryCatch(
      {
        FeaturePlot(Endo_ScAnemia, features = i, split.by = "group")
      },
      error = function(e) {
        # Handle error and print message
        print(paste("Error while creating FeaturePlot for gene:", i))
        print(e)
        dev.off()
        NULL
      }
    )
    # Print and save the plot if successful
  if (!is.null(p)) {
      print(p)
  } else {
      next
    }
  dev.off()
  }



sub_endo <- unique(Endo_ScAnemia$annot1)
  
  #dotplot WORKED
sub_endo <- unique(Endo_ScAnemia$annot1)
for (i in genes){
  for(j in sub_endo){
  f <- paste0("dotplots/Endothelial_subset_dotplots/", j, "_", i, ".png")
  sub_seurat <- subset(Endo_ScAnemia, annot1 %in% j)
  png(f, width = 7, height = 5, units = "in", res = 300)
  p <- DotPlot(sub_seurat, features = i, group.by = "group1")
  print(p)
  dev.off()
  }
}




# THSI WORKED WE SUBSET ALL CELLS IN FEATURE PLOT



sub_endo <- unique(Endo_ScAnemia$annot1)

# Loop through each gene and subgroup
for (i in genes) {
  for (j in sub_endo) {
    # Construct the file path for saving the plot
    f <- paste0("featureplots/Endothelial_subset_featureplots/", j, "_", i, ".png")
    
    # Subset the Seurat object by subgroup
    sub_seurat <- subset(Endo_ScAnemia, annot1 %in% j)

    # Check if subset is empty before proceeding
    if (ncol(sub_seurat) == 0) {
      print(paste("Skipping empty subset for group:", j))
      next  # Skip to the next iteration if the subset is empty
    }

    # Check for missing values in the feature data
    if (any(is.na(FetchData(sub_seurat, vars = i)))) {
      print(paste("Skipping gene:", i, "in group:", j, "due to missing values."))
      next  # Skip to the next iteration if there are missing values in the data
    }

    # Open PNG device to save the plot
    png(f, width = 19, height = 5, units = "in", res = 300)

    # Create the FeaturePlot for the specific gene in the subset
    p <- tryCatch(
      {
        FeaturePlot(sub_seurat, features = i, split.by = "group")
      },
      error = function(e) {
        # Handle error and print message
        print(paste("Error while creating FeaturePlot for gene:", i, "in group:", j))
        print(e)
        NULL
      }
    )

    # Print and save the plot if successful
    if (!is.null(p)) {
      print(p)
    }

    # Close the PNG device
    dev.off()
  }
}
























 
 
 getwd() 
  
png("dene.png", width = 7, height = 5, units = "in", res = 300)
 DotPlot(sickle_cell, features = "Pecam1", group.by = "group")
dev.off()

 
unique(factor(sickle_cell$group))




#Constructing the group name with ifelse
Endo_ScAnemia$group <- ifelse(grepl("_1$", colnames(Endo_ScAnemia)), "WT_Male",
                               ifelse(grepl("_2$", colnames(Endo_ScAnemia)), "WT Female",
                                      ifelse(grepl("_3$", colnames(Endo_ScAnemia)), "ScAnemia Male", "ScAnemia Female")))

Endo_ScAnemia$group1 <- Endo_ScAnemia$group
Endo_ScAnemia$group <- factor(Endo_ScAnemia$group, levels = c("WT_Male", "WT Female", "ScAnemia Male", "ScAnemia Female"))
Endo_ScAnemia$group1 <- factor(Endo_ScAnemia$group1, levels = rev(c("WT_Male", "WT Female", "ScAnemia Male", "ScAnemia Female")))
  
?factor


  
#IFNA1 =Ifnar1
#"Col4a3bp" = Cert

rownames(sickle_cell)[grep("Ifn", rownames(sickle_cell))]

FeaturePlot(Endo_ScAnemia, "Fabp4")

DimPlot(Endo_ScAnemia, group.by = "annot1")

```

#PATHWAY ANALYSIS- ELHAMDULILLAH!!!!!!!!!!!!!!!!!!!!!!!!!!

```{r}

library(org.Mm.eg.db)

sickle_cell <- readRDS("C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/SICKLE_PROJECT/sickle_cell_anemia_integrated_1_14_25.Rds")

#Integration after RNA LogNormalize of ScAnemia (4 samples)Vcp1, WT, Vcp2, WT-Hypoxia, Vcp3, ScAnemia-WT, Vcp4, ScAnemia-Hypoxia

WT_Male <- subset(sickle_cell, group %in% "WT_Male")
WT Female  <- subset(sickle_cell, group %in% "WT Female")
exp_norm <- as.matrix(WT_Male@assays[["RNA"]]@layers[["data"]])
exp_hyp <- as.matrix(WT Female@assays[["RNA"]]@layers[["data"]])


ref <- rowMeans(exp_norm)
sample <- rowMeans(exp_hyp)
rownames <- rownames(sickle_cell)
names(ref) <- rownames
names(sample) <- rownames

comb <- cbind(ref, sample)
comb <- as.data.frame(comb)
comb$Entrezid <- NA

# Convert gene symbols to Entrez IDs
gene_symbols <- rownames(comb)  # Assuming these are your gene symbols
entrez_ids <- mapIds(org.Mm.eg.db, 
                     keys = gene_symbols, 
                     column = "ENTREZID", 
                     keytype = "SYMBOL", 
                     multiVals = "first")

comb$Entrezid <- entrez_ids

filtered_comb <- comb[!is.na(comb$Entrezid), ]
rownames(filtered_comb) <- filtered_comb$Entrezid

filtered_comb <- as.data.frame(filtered_comb)
filtered_comb <- as.matrix(filtered_comb[ , c("ref", "sample")])
rownames <- rownames(filtered_comb)
#change to numeric forcefully
filtered_comb <- apply(filtered_comb, 2, as.numeric)
rownames(filtered_comb) <- rownames
any(!sapply(filtered_comb, is.numeric))  # Should return FALSE

library(gageData)
data(kegg.sets.mm)

gage_results <- gage(exprs = filtered_comb, gsets = kegg.sets.mm)

# View results
print(gage_results$greater)  # Up-regulated pathways
print(gage_results$less)     #

View(gage_results$greater)

write.csv(gage_results$greater, "with_means_gage.csv")




```



#TRY TO USE EXPRESSION MATRIXES -NULL

```{r}
rm(list = ls())

library(org.Mm.eg.db)
library(gageData)

sickle_cell <-  readRDS("C:/Users/skacar/OneDrive - Indiana University/PULMONARY POST DOC/___ScAnemia_PROJECT____/sickle_cell_11_15_24.Rds") #Integration after RNA LogNormalize of ScAnemia (4 samples)Vcp1, WT, Vcp2, WT-Hypoxia, Vcp3, ScAnemia-WT, Vcp4, ScAnemia-Hypoxia

WT_Male <- subset(sickle_cell, group %in% "WT_Male")
WT Female  <- subset(sickle_cell, group %in% "WT Female")
exp_norm <- as.matrix(WT_Male@assays[["RNA"]]@layers[["data"]])
exp_hyp <- as.matrix(WT Female@assays[["RNA"]]@layers[["data"]])
#ADD COL AND ROW NAMES CELLS AND GENES I MEAN
colnames(exp_norm) <- colnames(WT_Male)
colnames(exp_hyp) <- colnames(WT Female)
rownames(exp_norm) <- rownames(WT_Male)
rownames(exp_hyp) <- rownames(WT Female)


# Convert gene symbols to Entrez IDs
gene_symbols <- rownames(exp_norm)  # Assuming these are your gene symbols
entrez_ids <- mapIds(org.Mm.eg.db, 
                     keys = gene_symbols, 
                     column = "ENTREZID", 
                     keytype = "SYMBOL", 
                     multiVals = "first")

# Identify rows with valid Entrez IDs
valid_rows <- !is.na(entrez_ids)
# Filter the matrices
exp_norm <- exp_norm[valid_rows, ]
exp_hyp <- exp_hyp[valid_rows, ]

entrez_ids <- entrez_ids[valid_rows]
rownames(exp_norm) <- entrez_ids
rownames(exp_hyp) <- entrez_ids

exprs <- cbind(exp_norm, exp_hyp)

# Define reference and sample columns
ref_cols <- 1:ncol(exp_norm)
samp_cols <- (ncol(exp_norm) + 1):ncol(exprs)



#CHOOSE AS PROFILES
# Aggregate by averaging across cells for each group
ref_profile <- rowMeans(exprs[, ref_cols])  # WT_Male
samp_profile <- rowMeans(exprs[, samp_cols])  # WT Female

# Create a pseudobulk matrix
exprs_pseudobulk <- cbind(ref_profile, samp_profile)

library(gageData)
data("kegg.mm.gs")
# Run GAGE on pseudobulk profiles
gage_results <- gage(exprs = exprs_pseudobulk, gsets = kegg.sets.mm, ref = 1, samp = 2, compare = "unpaired")




class(gage_results$greater
write.csv(gage_results$greater, "all_Cells_group_gage.csv")


```




#USE DEGS instead of NORMZALIZED MATRIX- ELHAMDULILLAH!!!!!

#WT Female vs WT_Male

```{r}
library(Seurat)
degs <- FindMarkers(sickle_cell, ident.1 = "WT Female", ident.2 = "WT_Male", assay = "RNA", group.by = "group")
significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]

library(org.Mm.eg.db)
entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
names(entrez_ids) <- rownames(significant_degs)

entrez_valid <- entrez_ids[!is.na(entrez_ids)]
significant_degs <- significant_degs[names(entrez_valid), ]


exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
rownames(exprs_degs) <- entrez_valid


library(gageData)
data("kegg.sets.mm")
gage_results <- gage(exprs = exprs_degs, gsets = kegg.sets.mm, ref = NULL, samp = NULL, compare = "unpaired")

p <- gage_results$stats
#r <- gage_results$less
P <- as.data.frame(p)

P$mmu <- NA
P$mmu <- rownames(P)
# Remove the first word (everything before the first space) from rownames
P$mmu <- sub(" .*", "", P$mmu)
rownames(P) <- sub("^[^ ]+ ", "", rownames(P))
P$pathways <- NA
P$pathways <- rownames(P)


P <- P[order(-P$stat.mean), ]
#P <- P[-1, ] EXTRACT ROWS AS NEEDED
top_df <- head(P, 20)

```



```{r}
library(ggplot2)

p <- ggplot(top_df, aes(x = reorder(pathways, stat.mean), y = stat.mean)) +
    geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
   labs(title = "WT_Male vs WT Female for Whole Lung", x = "Pathways", y = "Mean Enrichment Score (stat.mean)") +
  theme_minimal()

png("PATHWAYS/WT_Male vs WT Female for Whole Lung_no_ribosome.png", width = 7, height = 5, units = "in", res = 300)
p
dev.off


```




#The other pairs

```{r}

 ident_pairs <- list(
    c("WT Female", "WT_Male"),
    c("ScAnemia Female", "WT Female"),
    c("ScAnemia Male", "WT_Male"),
    c("ScAnemia Female", "ScAnemia Male")
    
```


#DO ALL PAIRWISE COMPARISONS GAGE KEGG
```{r}

ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(ScAnemia_perismc, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(gageData)
  data("kegg.sets.mm")
  gage_results <- gage(exprs = exprs_degs, gsets = kegg.sets.mm, ref = NULL, samp = NULL, compare = "unpaired")
  p <- gage_results$stats
  #r <- gage_results$less
  P <- as.data.frame(p)
  P$mmu <- NA
  P$mmu <- rownames(P)
  # Remove the first word (everything before the first space) from rownames
  P$mmu <- sub(" .*", "", P$mmu)
  rownames(P) <- sub("^[^ ]+ ", "", rownames(P))
  P$pathways <- NA
  P$pathways <- rownames(P)
  P <- P[order(-P$stat.mean), ]
  if ("Ribosome" %in% rownames(P[1:20, ])) {
  # Remove the row corresponding to "Ribosome"
  P <- P[rownames(P) != "Ribosome", ]
}
  top_df <- head(P, 20)
  library(ggplot2)
  plot <- ggplot(top_df, aes(x = reorder(pathways, stat.mean), y = stat.mean)) +
    geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
   labs(title = paste0(ident_2, "_vs_", ident_1, " for Pericyte_Smc"), x = "Pathways", y = "Mean Enrichment Score (stat.mean)") +
  theme_minimal()
  png(paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, " for Pericyte_Smc.png"), width= 7, height = 5, units = "in", res = 300)
  plot
  print(plot)
  dev.off()
}
```



```{r}

library(ggplot2)

p <- ggplot(top_df, aes(x = reorder(pathways, stat.mean), y = stat.mean)) +
    geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
   labs(title = paste0(ident_2, "_vs_", ident_1, " for Whole Lung"), x = "Pathways", y = "Mean Enrichment Score (stat.mean)") +
  theme_minimal()

png(paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, " for Whole Lung.png"), width = 7, height = 5, units = "in", res = 300)
p
dev.off()

```






#from gage results u can continue
```{r, fig.height=4, fig.width=10}


ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(sickle_cell, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(ReactomePA)
  library(enrichplot)
  pathway_results <- enrichPathway(gene = rownames(exprs_degs), organism = "mouse", pvalueCutoff = 0.05)
  #ADDING LOG VALUES TO PATHWAY ANALYSIS
  pathway_results@result$logFC <- sapply(pathway_results@result$geneID, function(genes) {
  # Split geneID into individual Entrez IDs
  entrez_ids <- unlist(strsplit(genes, "/"))
  exprs_degs <- as.data.frame(exprs_degs)
  # Extract logFC values from exprs (column V1)
  logFC_values <- exprs_degs[entrez_ids, "V1"]
  
  # Compute the mean logFC for the pathway
  mean(logFC_values, na.rm = TRUE)
})
  pathway_results <- pairwise_termsim(pathway_results)
  #barplot(pathway_results, showCategory = 20)
  png(paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, "_dotplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- dotplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  png(paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, "_emapplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- emapplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  #emapplot(pathway_results, showCategory = 20, color = "logFC")  # Use logFC for node colors

}
```






```{r}
library(clusterProfiler)
library(org.Mm.eg.db)  # Mouse-specific annotation database


# Perform GO enrichment analysis
go_results <- enrichGO(
  gene          = rownames(exprs_degs),  # Use Entrez IDs as input genes
  OrgDb         = org.Mm.eg.db,          # Organism database for mouse
  keyType       = "ENTREZID",            # Specify the type of gene ID
  ont           = "BP",                  # Ontology type: BP, CC, or MF
  pAdjustMethod = "BH",                  # Adjust p-values using Benjamini-Hochberg
  pvalueCutoff  = 0.05,                  # Significance threshold for p-values
  qvalueCutoff  = 0.2,                   # Significance threshold for q-values
  readable      = TRUE                   # Converts Entrez IDs to gene symbols
)

# View the enrichment results
head(as.data.frame(go_results))

barplot(go_results, showCategory = 10, title = "Top 10 Enriched GO Biological Processes")

go_results@result



```







                       
 # ENRICH RESULT
 
```{r}
# Assuming your DEG list is called "deg_genes" and your reference list is called "reference_genes"
library(clusterProfiler)

# Example using enrichGO function
enrich_result <- enrichGO(
    gene          = mouse_data$X,        # Your DEG list
    OrgDb         = org.Mm.eg.db,     # Organism database (e.g., human genes)
    keyType       = "SYMBOL",         # Type of gene identifier (e.g., SYMBOL, ENSEMBL)
    ont           = "BP",             # Ontology type: BP, MF, or CC
    pAdjustMethod = "BH",             # p-value adjustment method
    universe      = rownames(sickle_cell),  # Custom reference list
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2
)
# View the enrichment results
head(as.data.frame(enrich_result))

barplot(enrich_result, showCategory = 10, title = "Top 10 Enriched GO Biological Processes")

```                      
                       
    

#Gaining ENRICHGO RESULTS AUTOMATICALLY PER PAIR

```{r}

ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(sickle_cell, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(clusterProfiler)
  library(org.Mm.eg.db)  # Mouse-specific annotation database
  # Perform GO enrichment analysis
  go_results <- enrichGO(
  gene          = rownames(exprs_degs),  # Use Entrez IDs as input genes
  OrgDb         = org.Mm.eg.db,          # Organism database for mouse
  keyType       = "ENTREZID",            # Specify the type of gene ID
  ont           = "BP",                  # Ontology type: BP, CC, or MF
  pAdjustMethod = "BH",                  # Adjust p-values using Benjamini-Hochberg
  pvalueCutoff  = 0.05,                  # Significance threshold for p-values
  qvalueCutoff  = 0.2,                   # Significance threshold for q-values
  readable      = TRUE                   # Converts Entrez IDs to gene symbols
)
  write.csv(go_results@result, paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, " enrichGo_results.csv"))
}
   
```
#KEGG RESULTS FOR PATHVIEW!!!

```{r}
kegg_result <- enrichKEGG(gene = rownames(exprs_degs),
                          organism = 'mmu',  # 'hsa' for Homo sapiens (change if necessary)
                          pvalueCutoff = 0.05)
head(kegg_result)

kegg_result_df <- as.data.frame(kegg_result)
#significant_results <- kegg_result_df[kegg_result_df$qvalue < 0.05, ]


#mmu05171

```




#AUTOMATIC KEGG RESULTS DF AND SAVING

```{r}

ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(sickle_cell, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  kegg_result <- enrichKEGG(gene = rownames(exprs_degs),
                          organism = 'mmu',  # 'hsa' for Homo sapiens (change if necessary)
                          pvalueCutoff = 0.05)
  kegg_result_df <- as.data.frame(kegg_result)
  #significant_results <- kegg_result_df[kegg_result_df$qvalue < 0.05, ]
  write.csv(go_results@result, paste0("./PATHWAYS/", ident_2, "_vs_", ident_1, " enrichKEGG_results.csv"))
}
   
```


#PATHVIEW

```{r}
library(pathview)
pv.out <- pathview(gene.data = exprs_degs,
                   pathway.id = "mmu05171",
                   species = "mmu",
                   out.suffix = "WT Female_vs_WT_Male",
                   kegg.native = TRUE)


pv.out <- pathview(gene.data = exprs_degs,
                   pathway.id = "mmu05171",
                   species = "mmu",
                   out.suffix = "WT Female_vs_WT_Male",
                   kegg.native = FALSE)
```











#KEGG PATHWAYS

```{r}

data("kegg.sets.mm")
#data("sigmet.idx.rn")
#kegg.sets.rn = kegg.sets.rn[sigmet.idx.rn]
kegg.sets.mm = kegg.sets.mm

```





# HOW TO CHECK THE DATA IN SPECIFIC PACKAGE AND CALL IT

```{r}
data()

data(package = "gage")

data(package = .packages(all.available = TRUE))

data("kegg_category", package = "clusterProfiler")


Data sets in package clusterProfiler:

DE_GSE8057                    Datasets gcSample contains a sample of gene clusters.
gcSample                      Datasets gcSample contains a sample of gene clusters.
kegg_category                 Datasets gcSample contains a sample of gene clusters.
kegg_species                  Datasets gcSample contains a sample of gene clusters.

length(unique(kegg_category$name))

rm(list = ls())
       
```



```{r}
library(KEGGREST)
# List all mouse pathways
mouse_pathways <- keggList("pathway", "mmu")
# Search for inflammation-related pathways
grep("inflammation|immune|cytokine|toll-like", mouse_pathways, value = TRUE)
grep("inflammation|toll|cytokine", mouse_pathways, value = TRUE)

df <- as.data.frame(mouse_pathways)


df$mouse_pathways <- sub(" - Mus musculus \\(house mouse\\)", "", df$mouse_pathways)

write.csv(df, "Keggrest_updated_pathways_mmu.csv")



```





```{r}
library(readxl)
library(AnnotationDbi)
library(org.Mm.eg.db)

# Add Entrez IDs to the original data
mouse_data$ENTREZID <- entrez_ids

# Extract fold changes and name them with Entrez IDs
foldchanges <- mouse_data$avg_log2FC
names(foldchanges) <- mouse_data$ENTREZID


keggres = gage(exprs = foldchanges, gsets = kegg.sets.mm, same.dir = FALSE)



# Check how many genes were mapped successfully
summary(is.na(mouse_data$ENTREZID))
#fold change value
summary(foldchanges)
# Count how many genes are used in pathway analysis
length(intersect(names(foldchanges), unlist(kegg.sets.mm)))

View(keggres$less)

write.csv(keggres$greater, "./gcapb_kegg_true.csv")



#EXTRACT GENES


# Check the gene set associated with the first pathway (Metabolic pathways, mo01100)
genes_in_metabolic_pathway <- kegg.sets.mm[["rno01100 Metabolic pathways"]]
metabolic_gcapb <- intersect(genes_in_metabolic_pathway, gcapb$ENTREZID)
class(metabolic_gcapb)
metabolic_gcapb <- as.data.frame(metabolic_gcapb)

metabolic_gcapb$gene_symbols <- NA
metabolic_gcapb$log_fold <- NA

matching_indices <- match(metabolic_gcapb$metabolic_gcapb, gcapb$ENTREZID)

metabolic_gcapb$log_fold <-  gcapb$gCapB.avg_log2FC[matching_indices]

# View the genes
print(genes_in_metabolic_pathway)


write.csv(metabolic_gcapb, "./metabolic_pathway_genes.csv")

```















#UMAPs

```{r, fig.height=19, fig.width=19}

#PERI SMOOTH
png("Pericyte_SMC_per_group.png", width = 19, height = 5, units = "in", res = 300)
DimPlot(ScAnemia_perismc, group.by = "Parse_StromAnn2", split.by = "group", pt.size = 2) + ggtitle("Pericytes and Smooth Muscle Cells")
dev.off()

png("Pericyte_SMC_all_together.png", width = 7, height = 5, units = "in", res = 300)
DimPlot(ScAnemia_perismc, group.by = "Parse_StromAnn2",  pt.size = 2, label = T) + NoLegend() + ggtitle("Pericytes and Smooth Muscle Cells")
dev.off()


DimPlot(ScAnemia_perismc, group.by = "Parse_StromAnn2", split.by = "group", pt.size = 2)
getwd()




png("Endothelial_cells_per_group.png", width = 19, height = 5, units = "in", res = 300)
DimPlot(Endo_ScAnemia, group.by = "annot1", split.by = "group", pt.size = 1) + ggtitle("Endothelial Cells")
dev.off()

png("Endothelial_cells_all_groups_together.png", width = 7, height = 5, units = "in", res = 300)
DimPlot(Endo_ScAnemia, group.by = "annot1",  pt.size = 1, label = T) + NoLegend() + ggtitle("Endothelial Cells")
dev.off()


DimPlot(sickle_cell, group.by = "Parse_StromAnn", split.by = "group", pt.size = 2)
getwd()


library(Seurat)
DimPlot(sickle_cell, group.by = "Parse_Ann1",  pt.size = 1, label = T) + NoLegend() + ggtitle("Whole Lung")


png("Whole_lung_all_groups_together.png", width = 13, height = 15, units = "in", res = 300)
DimPlot(sickle_cell, group.by = "Parse_Ann1",  pt.size = 1, label = T, repel = T) + NoLegend() + ggtitle("Whole_Lung")
dev.off()



png("Whole_lung_per_group.png", width = 30, height = 15, units = "in", res = 300)
DimPlot(sickle_cell, group.by = "Parse_Ann1",  pt.size = 1, label = T, repel = T, split.by = "group") + NoLegend() + ggtitle("Whole_Lung")
dev.off()


sickle_cell$Parse_Ann1 <-sickle_cell$Parse_Ann 
sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% "AT1"] <- "Mesothelial cells"
sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% c("gCapB", "gCapA", "gCapC", "gCapD", "gCapE")] <- "gCaps"
sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% c("Vascular Smc", "Airway Smc")] <- "Smooth muscle cells"
sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% "Aerocyte II"] <- "gCaps"
sickle_cell$Parse_Ann1[sub] <- "Aerocyte I"
sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% "Aerocyte I"] <- "aCaps"
sickle_cell$Parse_Ann2 <-sickle_cell$Parse_Ann1

sickle_cell$Parse_Ann1[sickle_cell$Parse_Ann1 %in% c("Vascular Smc", "Airway Smc")] <- "Smooth muscle cells"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("Peribronchial fibroblasts", "Alveolar fibroblasts", "Adventitial fibroblasts", "Myofibroblasts")] <- "Fibroblasts"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("Monocyte-derived Macrophages", "Interstitial Macrophages")] <- "Macrophages"

sickle_cell$Parse_Ann2 <- gsub(" proliferating$", "", sickle_cell$Parse_Ann2)
# Remove "Prolif. " from the beginning of cell names in the Parse_Ann2 column
sickle_cell$Parse_Ann2 <- gsub("^Prolif\\. ", "", sickle_cell$Parse_Ann2)

sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Act. Alveolar M", "Alveolar M")] <- "Alveolar Macrophages"

sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("Cd8+ NK cells", "Immature Cd8+ NK cells")] <- "NK cells"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("DC1", "DC2", "Plasmacytoid DCs")] <- "Dendritic cells"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("aCaps", "gCaps", "PAECs", "PVECs", "Lymphatic EC")] <- "Endothelial cells"

sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Th1 cells", "Th17 cells", "CD8+ T Cells", "Helper T Cells", "Tregs")] <- "T cells"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("AT1")] <- "Alveolar type I cell"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("AT2")] <- "Alveolar type II cell"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann1 %in% c("Transitional Pericytes")] <- "Smooth muscle cells"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Brush cells")] <- "Alveolar Macrophages"
sickle_cell$Parse_Ann2[sickle_cell$seurat_clusters %in% 6] <- "Alveolar Macrophages"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Mast cells")] <- NA
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Plasma cells")] <- NA
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Macrophages")] <- "Interstititonal Macrophages"
sickle_cell$Parse_Ann2[sickle_cell$Parse_Ann2 %in% c("Interstititonal Macrophages")] <- "Interstitial Macrophages"


DimPlot(sickle_cell, group.by = "Parse_Ann2",  pt.size = 1, label = T) + NoLegend() + ggtitle("Whole Lung")


sub <- subset(sickle_cell, umap_2 < -7 & umap_1 > 8)
sub <- colnames(sub)
DimPlot(sub)
table(sickle_cell$Parse_Ann2)


#Alveolar Macrophage Cd68, Mertk, Itgax(cd11c), siglec1
#interstitial Macrophage Adgr1, Itgam, (mrc1), Cd163, Cd14
Macrophages <- Cd68, Mertk, Adgre1(F4/80)
Macrophages <- c("Cd68", "Mertk", "Adgre1", "Marco", "Itgax", "Itgam", "Siglec1", "RT1-Da", "Mrc1", "Cd163", "Ccr2")   #Adgre1(F4/80) #RT1-Da-HLA_DRA

FeaturePlot(sickle_cell, Neutrophil_markers)
Neutrophil_markers <- c("S100a9", "S100a8", "Cxcr2", "Prtn3", "Sell", "Elane", "Il1b", "Stfa2", "Stfa2l1")


SCT_ScAnemia$Parse_Ann2 <- NA

SCT_ScAnemia$Parse_Ann2 <- sickle_cell$Parse_Ann2


DimPlot(sickle_cell, label = T)

DimPlot(sickle_cell, group.by = "Parse_Ann2",  pt.size = 1, label = T, repel = T) + NoLegend() + ggtitle("Whole Lung")

DimPlot(sickle_cell, cells.highlight = colnames(SCT_ScAnemia)[SCT_ScAnemia$Parse_Ann2 %in% "Neutrophils"]) + NoLegend() + ggtitle("Whole Lung")



DimPlot(sickle_cell, group.by = "Parse_Ann2",  pt.size = 1, label = T, repel = T) + NoLegend() + ggtitle("Whole Lung")
FeaturePlot(sickle_cell, "Mki67")
```


# MAKING TABLE FOR NEW

```{r}
# MAKING THE TABLE

sickle_cell$group1 <- sickle_cell$group1
table(sickle_cell$group1)
df <- as.data.frame(table(sickle_cell$Parse_Ann2))

df1 <- as.data.frame(table(sickle_cell$Parse_Ann2[sickle_cell$group1 == "WT_Male"]))
df2 <- as.data.frame(table(sickle_cell$Parse_Ann2[sickle_cell$group1 == "WT Female"]))
df3 <-as.data.frame(table(sickle_cell$Parse_Ann2[sickle_cell$group1 == "ScAnemia Male"]))
df4 <-as.data.frame(table(sickle_cell$Parse_Ann2[sickle_cell$group1 == "ScAnemia Female"]))

rownames(df) <- df$Var1

rownames(df1) <- df1$Var1
rownames(df2) <- df2$Var1
rownames(df3) <- df3$Var1
rownames(df4) <- df4$Var1

df$Var1 <- NULL
df1$Var1 <- NULL
df2$Var1 <- NULL
df3$Var1 <- NULL
df4$Var1 <- NULL

colnames(df)[1] <- "All_groups"
colnames(df1)[1] <- "WT Normoxia"
colnames(df2)[1] <- "WT Hypoxia"
colnames(df3)[1] <- "ScAnemia Normoxia"
colnames(df4)[1] <- "ScAnemia Hypoxia"





# Match rownames of df to rownames of df1
matched_indices <- match(rownames(df), rownames(df1))
df$WT_Male<- NA
df$WT_Male<- 0
df$WT_Male[!is.na(matched_indices)] <- df1[matched_indices[!is.na(matched_indices)], "WT Normoxia"]

matched_indices <- match(rownames(df), rownames(df2))
df$WT Female<- NA
df$WT Female<- 0
df$WT Female[!is.na(matched_indices)] <- df2[matched_indices[!is.na(matched_indices)], "WT Hypoxia"]

matched_indices <- match(rownames(df), rownames(df3))
df$ScAnemia Male<- NA
df$ScAnemia Male<- 0
df$ScAnemia Male[!is.na(matched_indices)] <- df3[matched_indices[!is.na(matched_indices)], "ScAnemia Normoxia"]

matched_indices <- match(rownames(df), rownames(df4))
df$ScAnemia Female<- NA
df$ScAnemia Female<- 0
df$ScAnemia Female[!is.na(matched_indices)] <- df4[matched_indices[!is.na(matched_indices)], "ScAnemia Hypoxia"]

df <- df[!rownames(df) %in% "AT1", ]
df <- df[!rownames(df) %in% "Basal Epithelial", ]

# ILLUSTRATING DF TABLE

# Load necessary libraries
library(ggplot2)
library(tidyr)

df$Cell_Type <- rownames(df)

# Convert data frame to long format
df_long <- pivot_longer(df, cols = c("WT_Male", "WT Female", "ScAnemia Male", "ScAnemia Female"),
                        names_to = "Condition", values_to = "Count")


library(gridExtra)
library(ggplot2) 
# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("Parse_Ann2_percentages.png", width = 15, height = 19, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()



# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = Count, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Comparison of Counts by Condition for Each Cell Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1") #+
  #scale_y_continuous(limits = c(0, 2500))


table(Endo_ScAnemia$orig.ident)
table(sickle_cell$orig.ident)

table(df$WT_Male)

#PERCENTAGES OF DF
df$WT_Male_per <- NA
df$WT_Male_per <- (df$WT_Male * 100) / 9211


df$WT Female_per <- NA
df$WT Female_per <- (df$WT Female * 100) / 8603

df$ScAnemia Male_per <- NA
df$ScAnemia Male_per <- (df$ScAnemia Male * 100) / 7427

df$ScAnemia Female_per <- NA
df$ScAnemia Female_per <- (df$ScAnemia Female * 100) / 7674

write.csv(df, "Parse_Ann2_df.csv")


#df$total <- NULL
#df$total[df_long$Condition == "WT_Male"] <- 9211
#df$total[df_long$Condition == "WT Female"] <- 8604
#df$total[df_long$Condition == "ScAnemia Male"] <- 7427
#df$total[df_long$Condition == "ScAnemia Female"] <- 7674



df_long$total <- NA
df_long$total[df_long$Condition == "WT_Male"] <- 9211
df_long$total[df_long$Condition == "WT Female"] <- 8603
df_long$total[df_long$Condition == "ScAnemia Male"] <- 7427
df_long$total[df_long$Condition == "ScAnemia Female"] <- 7674


df_long$count_percent <- NA
df_long$count_percent <- (df_long$Count * 100) / df_long$total


table(sickle_cell$orig.ident)

sum(df$ScAnemia Female)

# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = count_percent, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f", count_percent)), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "count_percentage", title = "Comparison of Percentages by Condition for Whole Lung Cell Types") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")

getwd()

yes <- rownames(sickle_cell)[grep("Stat", rownames(sickle_cell))]
class(yes)


for(i in yes){
  p <- DotPlot(sickle_cell, i, group.by = "group", scale.min = 0) + coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p)
}

for(i in genes){
  p <- DotPlot(sickle_cell, i, group.by = "group", scale.min = 0) + coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p)
}

colSums(df[, sapply(df, is.numeric)])
```

#MAKING TABLE FOR EC SUBTYPES

```{r}

# MAKING THE TABLE

Endo_ScAnemia$group1 <- sickle_cell$group1
table(Endo_ScAnemia$group1)
df <- as.data.frame(table(Endo_ScAnemia$annot1))

df1 <- as.data.frame(table(Endo_ScAnemia$annot1[Endo_ScAnemia$group1 == "WT_Male"]))
df2 <- as.data.frame(table(Endo_ScAnemia$annot1[Endo_ScAnemia$group1 == "WT Female"]))
df3 <-as.data.frame(table(Endo_ScAnemia$annot1[Endo_ScAnemia$group1 == "ScAnemia Male"]))
df4 <-as.data.frame(table(Endo_ScAnemia$annot1[Endo_ScAnemia$group1 == "ScAnemia Female"]))

rownames(df) <- df$Var1

rownames(df1) <- df1$Var1
rownames(df2) <- df2$Var1
rownames(df3) <- df3$Var1
rownames(df4) <- df4$Var1

df$Var1 <- NULL
df1$Var1 <- NULL
df2$Var1 <- NULL
df3$Var1 <- NULL
df4$Var1 <- NULL

colnames(df)[1] <- "All_groups"
colnames(df1)[1] <- "WT Normoxia"
colnames(df2)[1] <- "WT Hypoxia"
colnames(df3)[1] <- "ScAnemia Normoxia"
colnames(df4)[1] <- "ScAnemia Hypoxia"





# Match rownames of df to rownames of df1
matched_indices <- match(rownames(df), rownames(df1))
df$WT_Male<- NA
df$WT_Male<- 0
df$WT_Male[!is.na(matched_indices)] <- df1[matched_indices[!is.na(matched_indices)], "WT Normoxia"]

matched_indices <- match(rownames(df), rownames(df2))
df$WT Female<- NA
df$WT Female<- 0
df$WT Female[!is.na(matched_indices)] <- df2[matched_indices[!is.na(matched_indices)], "WT Hypoxia"]

matched_indices <- match(rownames(df), rownames(df3))
df$ScAnemia Male<- NA
df$ScAnemia Male<- 0
df$ScAnemia Male[!is.na(matched_indices)] <- df3[matched_indices[!is.na(matched_indices)], "ScAnemia Normoxia"]

matched_indices <- match(rownames(df), rownames(df4))
df$ScAnemia Female<- NA
df$ScAnemia Female<- 0
df$ScAnemia Female[!is.na(matched_indices)] <- df4[matched_indices[!is.na(matched_indices)], "ScAnemia Hypoxia"]


# ILLUSTRATING DF TABLE

# Load necessary libraries
library(ggplot2)
library(tidyr)

df$Cell_Type <- rownames(df)

# Convert data frame to long format
df_long <- pivot_longer(df, cols = c("WT_Male", "WT Female", "ScAnemia Male", "ScAnemia Female"),
                        names_to = "Condition", values_to = "Count")



colSums(df[, sapply(df, is.numeric)])


# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = Count, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = Count), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "Count", title = "Comparison of Counts by Condition for Each EC Subpopulations") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1") #+
  #scale_y_continuous(limits = c(0, 2500))


table(Endo_ScAnemia$orig.ident)
table(sickle_cell$orig.ident)

table(df$WT_Male)

#PERCENTAGES OF DF
df$WT_Male_per <- NA
df$WT_Male_per <- (df$WT_Male * 100) / 2730


df$WT Female_per <- NA
df$WT Female_per <- (df$WT Female * 100) / 1673

df$ScAnemia Male_per <- NA
df$ScAnemia Male_per <- (df$ScAnemia Male * 100) / 2306

df$ScAnemia Female_per <- NA
df$ScAnemia Female_per <- (df$ScAnemia Female * 100) / 2577

write.csv(df, "Endothelial_df.csv")


df_long$total <- NA
df_long$total[df_long$Condition == "WT_Male"] <- 2730
df_long$total[df_long$Condition == "WT Female"] <- 1673
df_long$total[df_long$Condition == "ScAnemia Male"] <- 2306
df_long$total[df_long$Condition == "ScAnemia Female"] <- 2577


df_long$count_percent <- NA
df_long$count_percent <- (df_long$Count * 100) / df_long$total



# Create grouped bar plot
p <- ggplot(df_long, aes(x = Cell_Type, y = count_percent, fill = Condition)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(label = sprintf("%.1f", count_percent)), 
          position = position_stack(vjust = 0.5), # Position the text labels in the middle of each bar section
          size = 3, color = "black") +
  labs(x = "Cell Type", y = "count_percent", title = "Comparison of count_percentages by Condition for Each EC Subpopulations") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set1")




library(gridExtra)
library(ggplot2) 
# Open a PNG device to save the output as a high-resolution image (300 dpi)
png("Endothelial_subpopulations_counts.png", width = 10, height = 15, units = "in", res = 300) #7

# Arrange and save all plots in a grid with 3 columns
p

# Close the graphics device
dev.off()
```


#----------------------------------------------------------------------
#----------------------------------------------------------------------
#----------------------------------------------------------------------
#----------------------------------------------------------------------
#----------------------------------------------------------------------
#Gaining ENRICHGO RESULTS PRE SUBGROUPS of EC cells

```{r}
Ec_subs <- table(Endo_ScAnemia$annot1)
Ec_subs <- names(Ec_subs)

for (i in Ec_subs){

  subcluster <- subset(Endo_ScAnemia, annot1 == i)
  
ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(subcluster, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(clusterProfiler)
  library(org.Mm.eg.db)  # Mouse-specific annotation database
  # Perform GO enrichment analysis
  go_results <- enrichGO(
  gene          = rownames(exprs_degs),  # Use Entrez IDs as input genes
  OrgDb         = org.Mm.eg.db,          # Organism database for mouse
  keyType       = "ENTREZID",            # Specify the type of gene ID
  ont           = "BP",                  # Ontology type: BP, CC, or MF
  pAdjustMethod = "BH",                  # Adjust p-values using Benjamini-Hochberg
  pvalueCutoff  = 0.05,                  # Significance threshold for p-values
  qvalueCutoff  = 0.2,                   # Significance threshold for q-values
  readable      = TRUE                   # Converts Entrez IDs to gene symbols
)
  write.csv(go_results@result, paste0("./PATHWAYS_new/", i, "_", ident_2, "_vs_", ident_1, " enrichGo_results.csv"))
}
}


   
```


#Gaining ENRICHGO RESULTS of SMC/pericytes and fibroblasts

```{r}
ScAnemia_subs <- table(sickle_cell$Parse_Ann2)
ScAnemia_subs <- names(ScAnemia_subs)
ScAnemia_subs <-ScAnemia_subs[c(10,17,18)] 


for (i in ScAnemia_subs){

  subcluster <- subset(sickle_cell, Parse_Ann2 == i)
  
ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(subcluster, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(clusterProfiler)
  library(org.Mm.eg.db)  # Mouse-specific annotation database
  # Perform GO enrichment analysis
  go_results <- enrichGO(
  gene          = rownames(exprs_degs),  # Use Entrez IDs as input genes
  OrgDb         = org.Mm.eg.db,          # Organism database for mouse
  keyType       = "ENTREZID",            # Specify the type of gene ID
  ont           = "BP",                  # Ontology type: BP, CC, or MF
  pAdjustMethod = "BH",                  # Adjust p-values using Benjamini-Hochberg
  pvalueCutoff  = 0.05,                  # Significance threshold for p-values
  qvalueCutoff  = 0.2,                   # Significance threshold for q-values
  readable      = TRUE                   # Converts Entrez IDs to gene symbols
)
  write.csv(go_results@result, paste0("./PATHWAYS_new/", i, "_", ident_2, "_vs_", ident_1, " enrichGo_results.csv"))
}
}


   
```

#REACTOME ANALYSIS for EC cells


```{r, fig.height=4, fig.width=10}

Ec_subs <- table(Endo_ScAnemia$annot1)
Ec_subs <- names(Ec_subs)

for (i in Ec_subs){

  subcluster <- subset(Endo_ScAnemia, annot1 == i)
  
  
ident_pairs <- list(
c("WT Female", "WT_Male"),
c("ScAnemia Female", "WT Female"),
c("ScAnemia Male", "WT_Male"),
c("ScAnemia Female", "ScAnemia Male"),
c("WT_Male", "WT Female"),
c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(subcluster, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(ReactomePA)
  library(enrichplot)
  pathway_results <- enrichPathway(gene = rownames(exprs_degs), organism = "mouse", pvalueCutoff = 0.05)
  #ADDING LOG VALUES TO PATHWAY ANALYSIS
  pathway_results@result$logFC <- sapply(pathway_results@result$geneID, function(genes) {
  # Split geneID into individual Entrez IDs
  entrez_ids <- unlist(strsplit(genes, "/"))
  exprs_degs <- as.data.frame(exprs_degs)
  # Extract logFC values from exprs (column V1)
  logFC_values <- exprs_degs[entrez_ids, "V1"]
  
  # Compute the mean logFC for the pathway
  mean(logFC_values, na.rm = TRUE)
})
  pathway_results <- pairwise_termsim(pathway_results)
  #barplot(pathway_results, showCategory = 20)
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_dotplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- dotplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_emapplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- emapplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  #emapplot(pathway_results, showCategory = 20, color = "logFC")  # Use logFC for node colors

}
}

```


```{r}

Ec_subs <- table(Endo_ScAnemia$annot1)
Ec_subs <- names(Ec_subs)
Ec_subs <- Ec_subs[4]

for (i in Ec_subs){

  subcluster <- subset(Endo_ScAnemia, annot1 == i)
  
  
ident_pairs <- list(
#c("WT Female", "WT_Male")#,
#c("ScAnemia Female", "WT Female")#, c("ScAnemia Male", "WT_Male")#, c("ScAnemia Female", "ScAnemia Male")#, c("WT_Male", "ScAnemia Male")#, c("ScAnemia Male", "ScAnemia Female")#, c("WT Female", "ScAnemia Female")#,c("WT_Male", "WT Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  write.csv(pair, "pair_current.csv")
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(subcluster, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(ReactomePA)
  library(enrichplot)
  pathway_results <- enrichPathway(gene = rownames(exprs_degs), organism = "mouse", pvalueCutoff = 0.05)

  #ADDING LOG VALUES TO PATHWAY ANALYSIS
  pathway_results@result$logFC <- sapply(pathway_results@result$geneID, function(genes) {
  # Split geneID into individual Entrez IDs
  entrez_ids <- unlist(strsplit(genes, "/"))
  exprs_degs <- as.data.frame(exprs_degs)
  # Extract logFC values from exprs (column V1)
  logFC_values <- exprs_degs[entrez_ids, "V1"]
  
  # Compute the mean logFC for the pathway
  mean(logFC_values, na.rm = TRUE)
})
  pathway_results <- pairwise_termsim(pathway_results)
        # Skip the rest if pairwise_termsim fails to calculate meaningful results
 
  #barplot(pathway_results, showCategory = 20)
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_dotplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- dotplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_emapplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- emapplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  #emapplot(pathway_results, showCategory = 20, color = "logFC")  # Use logFC for node colors

}
}

```

# percyte and smooth REACTOME
```{r}
ScAnemia_subs <- table(sickle_cell$Parse_Ann2)
ScAnemia_subs <- names(ScAnemia_subs)
ScAnemia_subs <-ScAnemia_subs[c(10,17,18)] 
ScAnemia_subs <-ScAnemia_subs[2] 


for (i in ScAnemia_subs){

  subcluster <- subset(sickle_cell, Parse_Ann2 == i)
  
  
ident_pairs <- list(
#c("WT Female", "WT_Male"),
#c("ScAnemia Female", "WT Female"),
#c("ScAnemia Male", "WT_Male"),
#c("ScAnemia Female", "ScAnemia Male"),
#c("WT_Male", "WT Female"),
#c("WT Female", "ScAnemia Female"),
c("WT_Male", "ScAnemia Male"),
c("ScAnemia Male", "ScAnemia Female")
)
        # Loop through each pair and find markers
for (pair in ident_pairs) {
  # Extract ident.1 and ident.2 from the pair
  ident_1 <- pair[1]
  ident_2 <- pair[2]
  library(Seurat)
  degs <- FindMarkers(subcluster, ident.1 = ident_2, ident.2 = ident_1, assay = "RNA", group.by = "group")
  significant_degs <- degs[degs$p_val_adj < 0.05 & (degs$pct.1 > 0.1 | degs$pct.2 > 0.1),  ]
  library(org.Mm.eg.db)
  entrez_ids <- mapIds(org.Mm.eg.db,
                     keys = rownames(significant_degs),
                     column = "ENTREZID",
                     keytype = "SYMBOL",
                     multiVals = "first")
  names(entrez_ids) <- rownames(significant_degs)
  entrez_valid <- entrez_ids[!is.na(entrez_ids)]
  significant_degs <- significant_degs[names(entrez_valid), ]
  exprs_degs <- matrix(significant_degs$avg_log2FC, ncol = 1)
  rownames(exprs_degs) <- entrez_valid
  
  library(ReactomePA)
  library(enrichplot)
  pathway_results <- enrichPathway(gene = rownames(exprs_degs), organism = "mouse", pvalueCutoff = 0.05)
  #ADDING LOG VALUES TO PATHWAY ANALYSIS
  pathway_results@result$logFC <- sapply(pathway_results@result$geneID, function(genes) {
  # Split geneID into individual Entrez IDs
  entrez_ids <- unlist(strsplit(genes, "/"))
  exprs_degs <- as.data.frame(exprs_degs)
  # Extract logFC values from exprs (column V1)
  logFC_values <- exprs_degs[entrez_ids, "V1"]
  
  # Compute the mean logFC for the pathway
  mean(logFC_values, na.rm = TRUE)
})
  pathway_results <- pairwise_termsim(pathway_results)
  #barplot(pathway_results, showCategory = 20)
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_dotplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- dotplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  png(paste0("./PATHWAYS_new/reactome", i, "_", ident_2, "_vs_", ident_1, "_emapplot_REACTOME.png"), width = 10, height = 15, res = 300, units = "in" )
  p <- emapplot(pathway_results, showCategory = 20)
  print(p)
  dev.off()
  #emapplot(pathway_results, showCategory = 20, color = "logFC")  # Use logFC for node colors

}
}

```



```{r, fig.height=10}
DimPlot(sickle_cell, group.by = "Parse_Ann1", label = T) +NoLegend()
DimPlot(Stroma_ScAnemia, group.by = "Parse_Ann", label = T) + NoLegend()

Stroma_ScAnemia$parse

FeaturePlot(sickle_cell, "Acta2")
FeaturePlot(sickle_cell, "Acta2")
c("Acta2", "Myh11", "Cnn1", "Tagln")

FeaturePlot(Stroma_ScAnemia, c("Acta2", "Myh11", "Cnn1", "Tagln"))

```




```{r, fig.height=10, fig.width=15}
SCMF_markers <- c("Agt", "Pdgfra", "Egfem1", "Htra1", "Tnc", "P2ry14", "Tgfbi", "Ltbp2", "Spon2")
SCMF_gpt <- c("Acta2", "Myh11", "Cnn1", "Col1a1", "Col3a1", "Pdgfra", "Fap", "Tagln", "Tpm2", "Sparc")
Pericyte_cell_card <- c("Lamc3", "Trpc6", "Cspg4", "Pdgfrb")
Myofibroblasts_gpt <- c("Acta2", "Tagln", "Myh11", "Tpm2", "Col1a1", "Col3a1", "Fn1", "Tnc", "Tgfb1", "Pdgfra", "Vcl", "Paxillin", "Des", "Vim", "Postn", "Spp1", "Thbs1")
Vascular_smo <- c("Map3k7cl","Aoc3", "Ntrk3","Gpc6")
Pnec <- c("Calca", "Grp", "Ascl1", "Scg5", "Cxc13", "Ngf", "Scg2", "Nov", 
           "Chgb", "Uchl1", "Pkib", "Ddc", "Snap25", "Pcsk1", "Snca", "Bex2", 
           "Dnajc12")


PMP <- c("Adh1", "Hoxb5", "Slc38a5", "Snai2", "Col13a1", "Ccnb1")
DimPlot(sickle_cell, group.by = "Parse_Ann1", label = T, repel = T) + NoLegend()

for(i in PMP){
    if(i %in% rownames(sickle_cell)){
  p <- FeaturePlot(sickle_cell, i, order = T, raster = F)
  print(p)
    }else {message(paste("Gene", i, "is", "missing", sep = " "))
    }
}


DimPlot(sickle_cell, group.by = "Parse_Ann1", split.by = "orig.ident", cells.highlight = colnames(sickle_cell)[grep("Myo", sickle_cell$Parse_Ann1)], raster = F)



Marker <- FindMarkers(sickle_cell, group.by = "Parse_Ann1", ident.1 = "Pericytes", only.pos = T, min.pct = 0.25, logfc.threshold = 0.25)
library(dplyr)
Marker <- Marker %>% 
  filter(p_val_adj < 0.5) %>%
  top_n(n =33, wt = avg_log2FC) #%>% 
  #filter(pct.1 > 0.700)#%>%
  #filter(pct.1 - pct.2 > 0.400)



Marker <- FindMarkers(sickle_cell, ident.1 = 21, only.pos = T, min.pct = 0.25, logfc.threshold = 0.25)
library(dplyr)
Marker <- Marker %>% 
  filter(p_val_adj < 0.5) %>%
  top_n(n =33, wt = avg_log2FC) #%>% 
  #filter(pct.1 > 0.700)#%>%
  #filter(pct.1 - pct.2 > 0.400)


FeaturePlot(sickle_cell, "Acta2", order = T, raster = F, split.by = "orig.ident")


DimPlot(sickle_cell, label = T, repel = T) + NoLegend()



write.csv(Sickle_AnnEC, "Sickle_AnnEC.csv")
write.csv(RNA__AnnEC, "RNA__AnnEC.csv")

```

# FINDING UNIQUES CORRESPONDENCE
```{r}
unique_correspondence <- RNA_ScAnemia@meta.data %>%
  group_by(Annotation_transfer_EC) %>%
  summarize(Parse_Ann2_correspondences = list(unique(Parse_Ann2)))


unique_correspondence <- as.data.frame(unique_correspondence)
class(unique_correspondence)

```


```{r}
# Create the mapping dataframe from the previous step
mapping_df <- unique_correspondence %>%
  tidyr::unnest(Parse_Ann2_correspondences) %>%
  rename(Annotation_transfer_EC = Annotation_transfer_EC, ParseAnn = Parse_Ann2_correspondences)

```





#sEX CHECK
```{r, fig.height=7, fig.width=6}

library(Seurat)
library(ggplot2)

mouse_sex_genes <- c(
  # Y-linked (Male-specific) genes
  "Ddx3y", "Eif2s3y", "Kdm5d", "Uty", "Zfy1", "Zfy2", "Sry", "Usp9y", 
  "Rbmy1a1", "Tspy1")
  
  mouse_sex_genes <- c(# X-linked genes (higher expression in females)
  "Xist", "Eif2s3x", "Kdm6a", "Rbm3", "Jarid1c", "Tspyl2", "Bcor", "Tmem29", 
  "Pir", "Eif1ax")
  
  mouse_sex_genes <- c(  # Autosomal genes with sex-biased expression
  "Foxl2", "Sox9", "Esr1", "Ar", "Dmrt1"
)


DotPlot(sickle_cell, mouse_sex_genes, group.by = "group", scale.min = 0) + coord_flip() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle(" Autosomal genes with sex-biased expression")



```

```{r}

anchors <- FindTransferAnchors(Par_smallref,
                               sickle_cell,
                               reference.assay = "RNA",
                               query.assay = "RNA"
                               )

Bis_Elh <- TransferData(anchors,
             as.list(Par_smallref@meta.data)
             )

# Loop through metadata columns and add to sickle_cell metadata

list <- names(Bis_Elh)

for(i in seq_along(list)){
  col_name <- list[i]  # Get metadata column name
  sickle_cell@meta.data[[col_name]] <- Bis_Elh[[col_name]]$predicted.id
}


```


```{r}
library(Matrix)

# Convert raw counts to a sparse matrix (reduces memory usage)
Parsall_comb@assays$RNA@layers$counts <- as(Parsall_comb@assays$RNA@layers$counts, "dgCMatrix")

# Save total UMI counts per cell in metadata
Parsall_comb@meta.data$total_UMIs <- Matrix::colSums(Parsall_comb@assays$RNA@layers$counts)

# Remove the full raw counts matrix
Parsall_comb@assays$RNA@layers$counts <- NULL


anchors <- FindTransferAnchors(Parsall_comb,
                               sickle_cell,
                               reference.assay = "RNA",
                               query.assay = "RNA"
                               )

Bis_Elh <- TransferData(anchors,
           Parsall_comb$Last_Call)
             )

Par_smallref@assays$RNA@layers$counts <- NULL

anchors <- FindTransferAnchors(Par_smallref,
                               sickle_cell,
                               reference.assay = "RNA",
                               query.assay = "RNA"
                               )

Bis_Elh <- TransferData(anchors,
           Par_smallref$Last_Call)
             


```
#NULLLLLLL_NOT WORKEDMAKING A SIMPLE REFERENCE SEURAT

```{r}

library(Seurat)

# Extract normalized data (already precomputed)
normalized_data <- Par_smallref@assays[["RNA"]]@layers[["data"]]

# Ensure row and column names are properly assigned
rownames(normalized_data) <- rownames(Par_smallref)
colnames(normalized_data) <- colnames(Par_smallref)

# Create Seurat object, storing in counts (Seurat expects raw counts)
reference_seurat <- CreateSeuratObject(counts = normalized_data)

# Manually assign pre-normalized data to the `data` slot
reference_seurat@assays$RNA@layers$data <- normalized_data
reference_seurat@assays$RNA@layers$counts <- normalized_data

# Assign metadata (cell labels) back
reference_seurat@meta.data <- Par_smallref@meta.data

reference_seurat <- FindVariableFeatures(reference_seurat)
#LETS USE
anchors <- FindTransferAnchors(reference_seurat,
                               sickle_cell,
                               reference.assay = "RNA",
                               query.assay = "RNA",
                                features = common_features
                               )

Bis_Elh <- TransferData(anchors,
           as.list(reference_seurat@meta.data))
             )
class(reference_seurat@assays$RNA)
Layers(reference_seurat)


```


```{r}
common_features <- intersect(rownames(reference_seurat), rownames(sickle_cell))
length(common_features)  # This should not be zero!

yes <- Seurat::FindTransferAnchors

```


```{r}
Idents(seurat_obj) <- seurat_obj@meta.data$group  # Assign groups
DEG_genotype <- FindMarkers(seurat_obj, ident.1 = c("VPC1", "VPC3"), ident.2 = c("VPC2", "VPC4"))
DEG_sex <- FindMarkers(seurat_obj, ident.1 = c("VPC1", "VPC2"), ident.2 = c("VPC3", "VPC4"))

```


```{r}
library(clusterProfiler)
gsea <- gseGO(geneList = sort(DEG_genotype$logFC, decreasing = TRUE), OrgDb = org.Hs.eg.db, ont = "BP")
dotplot(gsea)

```



```{r}
3 Compare cluster separation (density-based approach)

r
Copy
Edit
library(fpc)
db <- cluster.stats(dist(Embeddings(seurat_obj[["umap"]])), clusters = Idents(seurat_obj))
db$avg.silwidth  # Measures cluster separation


 Better Approach: Overlay All Samples in One UMAP
Instead of four separate UMAPs, you can combine all samples and color them by group:

r
Copy
Edit
seurat_obj$sample <- factor(c(rep("VPC1", n1), rep("VPC2", n2), rep("VPC3", n3), rep("VPC4", n4)))
DimPlot(seurat_obj, group.by = "sample")
```



```{r, fig.width=21, fig.height=8}

DimPlot(RNA_ScAnemia, split.by  = "group")



```



